---
title: "Class and mortality in NHIS"
author: "Jerzy Eisenberg-Guyot"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    keep_md: true
always_allow_html: true
---

* Analysis:

    + subdivide workers (and managers?) based on ONET or Krieger's method
    + compare descriptives and results using classwk/classwk2 2001-2003 when the variables overlap
    + incorporate firmno variable (numemps) for capitalist vs PB from 1997-2014 as sensitivity/descriptive analysis
    + sensitivity using survey design object for coxph - show SEs are similar to main
    + estimate proportions in classes over time
    + estimate class*year interaction term model as sensitivity
    
* Investigate: 
    + need to handle hispanic oversample correctly: advice here: https://nhis.ipums.org/nhis/userNotes_Hispanic_oversample.shtml
        + do so by using inverse probability of selection weights in sensitivity analyses? Tig says that makes sense, check if anyone ever responds on IPUMS forum

* General notes:
    + originally was hoping to examine cause-specific mortality, but I'm not sure it makes sense as counts are small, detailed cause of death not available from 2005-on, and would need to handle competing events
    + i don't think it's worth doing MICE since there's very little missingness and it'd be incredibly computationally intensive given the size of the dataset

* Methods notes: 
    + when robust=T, weights are treated properly by survfit and coxph functions (as probability or sampling weights rather than frequency weights)
    + citations for propensity score approach: https://www.ncbi.nlm.nih.gov/labs/pmc/articles/PMC5802372/; https://journals.sagepub.com/doi/full/10.1177/0193841X20938497
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(data.table)
library(here)
library(survival)
library(rms)
library(broom)
library(Rcpp)
library(rstpm2)
library(survey)
library(kableExtra)
library(tableone)
library(survminer)
library(ipw)
library(directlabels)
library(RColorBrewer)
library(patchwork)
options(scipen=999)
options(knitr.kable.NA = '')

#load dat
dat <- fread(here('nhis_1986_2014.csv'))

#make all variable names lowercase
dat %>%
  rename_all(tolower) -> dat

#exclude those outside age ranges (age has 0.09% missing)
#exclude those ineligible for mortality follow-up and those <25 or 65+ and those who aren't sample adults from 1997-on (occ1995 and classwk not available for others after that time)
#exclude 1997-2000, when there's no data on whether business is incorporated or not
#exclude 2015-on, when there's no mortality data
#exclude 1992 Hispanic oversample (only 4641 additional respondents) per advice of IPUMS
dat %>%
  filter((age>=25 & age<65) & mortelig==1 & (is.na(astatflg) | astatflg==1) & (year<1997 | (year>2000 & year<2015)) & !(year==1992 & substr(nhispid, 1, 4)==1991)) -> dat_sub

#make variables
dat_sub %>%
  mutate_at(.vars=(vars(-c(mortucod, mortwt, mortwtsa, psu, strata))), ~ifelse(. %in% c(91, 96, 97, 98, 99, 970, 980, 990, 7777, 8888, 9999), NA, .)) %>%
  mutate(classwk2=ifelse(classwk2 %in% (7:9), NA, classwk2), #these are real values in some other variables so we'll reset them here
         racesr=ifelse(racesr==900, NA, racesr)) %>%
  mutate(year_1986=year-1986, #center year at beginning and end of follow-up
         year_2014=year-2014,
         mortwt_f=ifelse(year<=1996, mortwt, mortwtsa), #per advice of IPUMS help desk, use mortwt from 1986-1996 and mortwtsa from 1997-on
         dead=ifelse(mortstat==1, 1, 0),
         mortdodq=ifelse(mortdodq==1, 91.25, #assume deaths occurred at the end of the quarter
                         ifelse(mortdodq==2, 182.5,
                                ifelse(mortdodq==3, 273.75, 
                                       ifelse(mortdodq==4, 365, NA)))),
         time=ifelse(dead==0, 2016 - year, mortdody + (mortdodq/365) - year), #deaths were allowed to occur through the end of Dec 31 2015, i.e., 2016
         dead_1996=ifelse(year>1996, NA,
                          ifelse(year<=1996 & (is.na(mortdody) | mortdody>1996), 0,
                                 ifelse(year<=1996 & mortdody<=1996, 1, NA))),
         time_1996=ifelse(year>1996, NA,
                          ifelse(dead_1996==1, mortdody + (mortdodq/365) - year,
                                 ifelse(dead_1996==0, 1997-year, NA))),
         manager=ifelse(occ1995>=102 & occ1995<=104, 1, 0),
         class=factor(ifelse(empstat==220, "NILF",
                             ifelse((empstat>=210 & empstat<=214) | (year<2001 & (classwk>=20 & classwk<=34) & manager!=1) | (year>=2001 & (classwk2>=1 & classwk2<=4) & manager!=1), "Worker",
                                    ifelse((year<2001 & (classwk>=20 & classwk<=34) & manager==1) | (year>=2001 & (classwk2>=1 & classwk2<=4) & manager==1), "Manager",
                                           ifelse((year<2001 & classwk==41) | (year>=2001 & classwk2==5 & jobsecincorp==2), "Capitalist",
                                                  ifelse((year<2001 & (classwk==42 | classwk==50)) | ((year>=2001 & classwk2==6) | (year>=2001 & classwk2==5 & jobsecincorp==1)), "PB", NA))))), 
                     levels=c("Capitalist", "PB", "Manager", "Worker", "NILF")),
         race=ifelse(racesr==100, "White", 
                     ifelse(racesr==200, "Black", "Other")),
         race_h=factor(ifelse(hispeth!=10, "Hispanic",
                              ifelse(hispeth==10 & race=="White", "NH white",
                                     ifelse(hispeth==10 & race=="Black", "NH black", "NH other"))), levels=c("NH white", "NH black", "Hispanic", "NH other")),
         poc=ifelse(race_h=="NH white", "NH white", "POC"),
         sex=ifelse(sex==1, "male", "female"),
         educ=factor(ifelse(educrec1<13, "<HS",
                            ifelse(educrec1==13, "HS",
                                   ifelse(educrec1==14, "Some college", "College+"))), levels=c("<HS", "HS", "Some college", "College+")),
         marital_tri=factor(ifelse(marstat %in% 10:12, "Married",
                                   ifelse(marstat %in% 20:40, "Widowed/divorced/separated", 
                                          ifelse(marstat==50, "Single", NA))), levels=c("Married", "Single", "Widowed/divorced/separated")),
         region=ifelse(region==1, "NE", 
                       ifelse(region==2, "MW",
                              ifelse(region==3, "S", "W"))),
         class_gender=factor(ifelse(class=="Worker" & sex=="male", "Male worker",
                                    ifelse(class=="Worker" & sex=="female", "Female worker",
                                           ifelse(class=="Manager" & sex=="male", "Male manager",
                                                  ifelse(class=="Manager" & sex=="female", "Female manager",
                                                         ifelse(class=="PB" & sex=="male", "Male PB",
                                                                ifelse(class=="PB" & sex=="female", "Female PB",
                                                                       ifelse(class=="Capitalist" & sex=="male", "Male capitalist",
                                                                              ifelse(class=="Capitalist" & sex=="female", "Female capitalist",
                                                                                     ifelse(class=="NILF" & sex=="male", "Male NILF",
                                                                                            ifelse(class=="NILF" & sex=="female", "Female NILF", NA)))))))))),
                             levels=c("Male capitalist", "Male PB", "Male manager", "Male worker", "Male NILF",
                                      "Female capitalist", "Female PB", "Female manager", "Female worker", "Female NILF")),
         class_poc=factor(ifelse(class=="Worker" & poc=="NH white", "NH white worker",
                                 ifelse(class=="Worker" & poc=="POC", "POC worker",
                                        ifelse(class=="Manager" & poc=="NH white", "NH white manager",
                                               ifelse(class=="Manager" & poc=="POC", "POC manager",
                                                      ifelse(class=="PB" & poc=="NH white", "NH white PB",
                                                             ifelse(class=="PB" & poc=="POC", "POC PB",
                                                                    ifelse(class=="Capitalist" & poc=="NH white", "NH white capitalist",
                                                                           ifelse(class=="Capitalist" & poc=="POC", "POC capitalist", 
                                                                                  ifelse(class=="NILF" & poc=="NH white", "NH white NILF", 
                                                                                         ifelse(class=="NILF" & poc=="POC", "POC NILF", NA)))))))))),
                          levels=c("NH white capitalist", "NH white PB", "NH white manager", "NH white worker", "NH white NILF",
                                   "POC capitalist", "POC PB", "POC manager", "POC worker", "POC NILF")),
         class_region=factor(ifelse(class=="Worker" & region=="MW", "MW worker",
                                    ifelse(class=="Worker" & region=="NE", "NE worker",
                                           ifelse(class=="Worker" & region=="S", "S worker",
                                                  ifelse(class=="Worker" & region=="W", "W worker",
                                                         ifelse(class=="Manager" & region=="MW", "MW manager",
                                                                ifelse(class=="Manager" & region=="NE", "NE manager",
                                                                       ifelse(class=="Manager" & region=="S", "S manager",
                                                                              ifelse(class=="Manager" & region=="W", "W manager",
                                                                                     ifelse(class=="PB" & region=="MW", "MW PB",
                                                                                            ifelse(class=="PB" & region=="NE", "NE PB",
                                                                                                   ifelse(class=="PB" & region=="S", "S PB",
                                                                                                          ifelse(class=="PB" & region=="W", "W PB",
                                                                                                                 ifelse(class=="Capitalist" & region=="MW", "MW capitalist",
                                                                                                                        ifelse(class=="Capitalist" & region=="NE", "NE capitalist",
                                                                                                                               ifelse(class=="Capitalist" & region=="S", "S capitalist",
                                                                                                                                      ifelse(class=="Capitalist" & region=="W", "W capitalist", 
                                                                                                                                             ifelse(class=="NILF" & region=="MW", "MW NILF",
                                                                                                                                                    ifelse(class=="NILF" & region=="NE", "NE NILF",
                                                                                                                                                           ifelse(class=="NILF" & region=="S", "S NILF",
                                                                                                                                                                  ifelse(class=="NILF" & region=="W", "W NILF", NA)))))))))))))))))))),
                             levels=c("MW capitalist", "NE capitalist", "S capitalist", "W capitalist",
                                      "MW PB", "NE PB", "S PB", "W PB", 
                                      "MW manager", "NE manager", "S manager", "W manager",
                                      "MW worker", "NE worker", "S worker", "W worker",
                                      "MW NILF", "NE NILF", "S NILF", "W NILF"))) -> dat_sub
                                                                                     
#exclude couple respondents with negative follow-up time
dat_sub %>%
  filter(time>=0) -> dat_sub
```

# Missingness 

Age missingness is in full dataset; missingness for other variables is among those ages 25-64

```{r message=FALSE, warning=FALSE}
#age in full dataset
dat %>%
  mutate_at(.vars=(vars(-c(mortucod, mortwt, mortwtsa, psu, strata))), ~ifelse(. %in% c(91, 96, 97, 98, 99, 970, 980, 990, 7777, 8888, 9999), NA, .)) %>%
  summarise_at(vars('age'), funs(na=100*sum(is.na(.)/1233801))) %>%
  mutate(name='age') %>% 
  relocate(na, .after=last_col()) -> dat_age_na

#other vars in subsetted dataset
dat_sub %>%
  summarise_at(vars(c('class', 'sex', 'race_h', 'educ', 'marital_tri', 'region')), funs(100*sum(is.na(.)/861724))) %>%
  tidyr::pivot_longer(class:region, values_to='na') -> dat_other_na

kable(bind_rows(dat_age_na, dat_other_na), digits=2, col.names=c("Variable", "Percent missing")) %>%
  kable_styling('striped')
```

# Descriptives

## Unweighted and stratified with missingness in variables of interest (aside from age and class)

```{r message=FALSE, warning=FALSE}
#vars of interest
vars <- c('sex', 'race_h', 'educ', 'marital_tri', 'region', 'age')
catvars <- c("sex")
nonorm <- c('age')

x <- CreateTableOne(data = dat_sub, vars=vars, factorVars=catvars, strata='class', includeNA=TRUE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p')
kable(x[,1:5]) %>%
  kable_styling(c("striped", "condensed"))
```

# Class and all-cause mortality

```{r}
#plotting function
plotted <- function(datted=kapped, caps=120, pbs=120, mans=120, works=120, nilfs=120){
  
  dat_surv <- data.frame(time=datted$time,
                         surved=datted$surv,
                         se=datted$std.err,
                         class=c(rep("Caps", caps), rep("PBs", pbs), rep("Mans", mans), rep("Wrks", works), rep("NILFs", nilfs)))

  ggplot(dat_surv, aes(x=time, y=surved, group=class, color=class, fill=class, label=class)) + 
    geom_line() +
    geom_ribbon(aes(ymin=surved-1.96*se, ymax=surved+1.96*se, color=class), alpha=0.2, lty=0) +
    geom_dl(method='last.qp') +
    scale_color_manual(values=brewer.pal(n = 12, name = "Paired")[c(2,4,6,8,10)]) +
    scale_fill_manual(values=brewer.pal(n = 12, name = "Paired")[c(2,4,6,8,10)]) +
    xlab("Years") +
    ylab("Probability of survival") +
    scale_x_continuous(expand=expansion(mult=c(0,0.09))) +
    scale_y_continuous(expand=expansion(mult=c(0.01,0))) +
    theme_light() +
    theme(legend.position="none")
}

#survdiff function
survdiffed <- function(datted=kapped, timed=30){
  
  #end of f/u
  point <- summary(datted,times=timed)
  
  #pb-cap
  pb <- 100*(point$surv[2] - point$surv[1])
  pb_lower <- 100*(pb/100 - 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  pb_upper <- 100*(pb/100 + 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
    
  #mc-cap
  mc <- 100*(point$surv[3] - point$surv[1])
  mc_lower <- 100*(mc/100 - 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  mc_upper <- 100*(mc/100 + 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  
  #worker-cap
  wc <- 100*(point$surv[4] - point$surv[1])
  wc_lower <- 100*(wc/100 - 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  wc_upper <- 100*(wc/100 + 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  
  #nilf-cap
  nc <- 100*(point$surv[5] - point$surv[1])
  nc_lower <- 100*(nc/100 - 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  nc_upper <- 100*(nc/100 + 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  
  rbind(c(pb, pb_lower, pb_upper), c(mc, mc_lower, mc_upper), c(wc, wc_lower, wc_upper), c(nc, nc_lower, nc_upper))
  
}

#tidying regression output function
tidy_n <- function(modded=mod, rows=1:4, cols=c(1,2,7,8), bind=binded, captioned="Ref: capitalist") {
  tidydf <- cbind(tidy(modded, exponentiate=T, conf.int=T)[rows, cols], bind)
  tidydf$N <- c(length(modded$residuals), NA, NA, NA)
  kable(tidydf, digits=c(2,2,2,2,1,1,1,1), caption=captioned, col.names=c(" ", "HR", "Lower", "Upper", "SD", "Lower", "Upper", "N")) %>%
    kable_styling("striped") %>%
    add_header_above(c(" ", "HR"=3, "SD per 100 at end of f/u"=3, " "))
}

#alternative KM approach using ggsurvplot - makes it look like more of a traditional KM plot (with steps), but is harder to read
#plotted <- ggsurvplot(kapped, 
#           ylim=c(0.6, 1.0),
#           conf.int = T,
#           axes.offset=T,
#           censor=F,
#           palette=brewer.pal(n = 12, name = "Paired")[c(2,4,6,8,10)],
#           legend="none",
#           ggtheme=theme_light()) +
#  ylab("Probability of survival") +
#  xlab("Year") 
#
#plotted %++%
#  geom_dl(aes(label = class), method = 'last.qp')
```

## Overall

### Distribution of IPW

```{r warning=FALSE, message=FALSE}
dat_sub %>%
  filter(!is.na(class) & !is.na(age) & !is.na(sex) & !is.na(year)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 5), 
                     data = subset(dat_sub, !is.na(class) & !is.na(age) & !is.na(sex) & !is.na(year)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_overall

summary(dat_sub_overall$sw)
```

```{r warning=FALSE, message=FALSE}
kapped <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_overall, se=T)
mod <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_overall)
binded <- survdiffed()
```

### Survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=6}
plotted()

ggsave("overall_survival.png", dpi=600, height=4, width=6)
```

### Survival difference and hazard ratio

```{r}
tidy_n()
```

## 1986-1995 waves w/ f/u through end of 1996 vs 2005-2014 waves w/ f/u through end of 2015

### Distribution of IPW

#### Early wave

```{r warning=FALSE, message=FALSE}
dat_sub %>%
  filter(year<=1995 & !is.na(class) & !is.na(age) & !is.na(sex) & !is.na(year)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 3), 
                     data = subset(dat_sub, year<=1995 & !is.na(class) & !is.na(age) & !is.na(sex) & !is.na(year)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_1986

summary(dat_sub_1986$sw)
```

#### Late wave

```{r warning=FALSE, message=FALSE}
dat_sub %>%
  filter(year>=2005 & !is.na(class) & !is.na(age) & !is.na(sex) & !is.na(year)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 3), 
                     data = subset(dat_sub, year>=2005 & !is.na(class) & !is.na(age) & !is.na(sex) & !is.na(year)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_2005

summary(dat_sub_2005$sw)
```

```{r warning=FALSE, message=FALSE}
#early
kapped_early <- survfit(Surv(time_1996, dead_1996)~class, robust=T, w=sw_f, data=dat_sub_1986, se=T)
mod_early <- coxph(Surv(time_1996, dead_1996)~class, robust=T, w=sw_f, data=dat_sub_1986)
binded_early <- survdiffed(kapped_early, timed=11)

#late
kapped_late <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_2005, se=T)
mod_late <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_2005)
binded_late <- survdiffed(kapped_late, timed=11)
```

### Survival plots

```{r warning=FALSE, message=FALSE, fig.height=4.5, fig.width=8.5}
plotted(datted=kapped_early, caps=43, pbs=44, mans=43, works=44, nilfs=44) + 
  ggtitle("1986-1995 with follow-up through 1996") + 
  scale_y_continuous(limits=c(0.88, 1.0), breaks=c(0.88, 0.91, 0.94, 0.97, 1.0), expand=expansion(mult=c(0,0))) +
  scale_x_continuous(breaks=seq(1, 11, 1), expand=expansion(mult=c(0,0.11))) +
plotted(datted=kapped_late, caps=33, pbs=44, mans=44, works=44, nilfs=44) + 
  ggtitle("2005-2014 with follow-up through 2015") + 
  scale_y_continuous(limits=c(0.88, 1.0), breaks=c(0.88, 0.91, 0.94, 0.97, 1.0), expand=expansion(mult=c(0,0))) +
  scale_x_continuous(breaks=seq(1, 11, 1), expand=expansion(mult=c(0,0.11))) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())

ggsave("change_over_time.png", dpi=600, height=4, width=8.5)
```

### Survival difference and hazard ratio

```{r}
tidy_n(modded=mod_early, bind=binded_early, captioned="Ref: capitalists (1986-1995 subset)")
tidy_n(modded=mod_late, bind=binded_late, captioned="Ref: capitalists (2005-2014 subset)")
```

# Session info

```{r  warning=FALSE, message=FALSE}
sessionInfo()
```
