---
title: "Dead labor: classed, gendered, and racialized mortality inequities in the United States"
author: "Jerzy Eisenberg-Guyot"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    keep_md: true
    toc_depth: 5
always_allow_html: true
---

* Overview: 
  + Descriptive analyses of class and mortality in the cross-sectional 1986-2018 NHIS with mortality follow-up through the end of 2019. Survival curves and Cox models are minimally adjusted (age, gender, and year) using IPW unless otherwise noted.
<br>
* Coding of classes:
  + Incorporated business owners (IBOs, loosely analogous to capitalists) are those who are self-employed in an incorporated business
  + Unincorporated business owners (UBOs, loosely analogous to the petit bourgeoisie) are those who are self-employed in an unincorporated business or working without pay in a family business/farm (very few people in the latter)
  + Managers (mans) are those who are not self-employed but who are employed and who have an "executive, administrative, or managerial" occupation
  + Workers (wrks) are those who are unemployed/"not employed" (few people) OR who are employed but do not have a management occupation
  + Not in the labor force (NLFs) are those who are "not in the labor force" 
<br>
* To do:
  + Make sure ok to link IPUMS NHIS to CDC NHIS (e.g., that IPUMS doesnt have concerns about data quality - check back into forum periodically)
<br>
* Useful references regarding IBOs vs UBOs:
  + Nuts and bolts: 
    - https://smallbusiness.chron.com/difference-between-incorporated-unincorporated-businesses-57463.html
    - https://www.upcounsel.com/is-an-llc-incorporated-or-unincorporated
    - https://www.uschamber.com/co/start/strategy/sole-proprietorship-vs-incorporation
    - https://portal.311.nyc.gov/article/?kanumber=KA-03067: states LLCs are unincorporated (also stated elsewhere online)
  + Demographics: 
    - https://www.pewresearch.org/social-trends/2015/10/22/three-in-ten-u-s-jobs-are-held-by-the-self-employed-and-the-workers-they-hire/
    - https://www.bls.gov/opub/mlr/2010/09/art2full.pdf
    - https://www.census.gov/content/dam/Census/library/working-papers/2018/demo/SEHSD-WP2018-28.pdf
<br>
* General notes:
  + Don't think it makes sense to incorporate  income into the regression analyses (e.g., mediation), even just as a sensitivity analysis. The 'decent' income data is only available from 1997-2018, and even that data is categorical and top-coded. Moreover, the top-coding changed over time (75k from 1997-2006; 100k from 2007-2008; and 115k from 2009-2018). Thus, unless we extrapolated or imputed beyond the top-codes, we'd have to use a categorical income variable with a max category of $75k+, which is well below the third quartile of family incomes after adjusting for inflation (and below IBOs' and mans' median incomes). The original income data also contains a lot of missingness, and NHIS recommends it be analyzed via imputation methods, which would complicate our approach considerably. 
  + Too few respondents in minoritized racial/ethnic groups who are IBOs to break out minoritized racial/ethnic groups into finer subgroups (survival estimates very unstable)
  + Numemps (number of employees at work) only available from 2001-2018 - would be alternative way to classify business owners but I think its categories are too imprecise and limited availability is problematic
  + Can't identify "chief executives, general and operative managers, and legislators" until 2004 because of limited occupation data before then (can never identify "chief executives" alone, who we've classified as IBOs in the past)
  + Useful citation regarding occupation and mortality in NHIS: https://www.sciencedirect.com/science/article/pii/S0091743520301535 and http://paa2019.populationassociation.org/uploads/190718
<br>
* Methods notes: 
  + When robust=T, weights are treated properly by survfit and coxph functions (as probability or sampling weights rather than frequency weights)
  + Citations for IPW and survey weighting approach: https://www.ncbi.nlm.nih.gov/labs/pmc/articles/PMC5802372/ and https://journals.sagepub.com/doi/full/10.1177/0193841X20938497
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(data.table)
library(here)
library(survival)
library(rms)
library(broom)
library(Rcpp)
library(survey)
library(kableExtra)
library(tableone)
library(survminer)
library(ipw)
library(directlabels)
library(RColorBrewer)
library(patchwork)
library(cobalt)
library(naniar)
library(ggmosaic)
library(multcomp)
options(scipen=999)
options(knitr.kable.NA = '')

#####manage ipums dat
#load ipums dat
dat_ipums <- fread(here('ipums_dat_1986_2018.csv'))

#make all variable names lowercase, fix merging id, and remove mortality variables
dat_ipums %>%
  rename_all(tolower) %>%
  mutate(nhispid=as.numeric(nhispid)) %>%
  dplyr::select(-c(mortelig, mortstat, mortdodq, mortdody, mortucod, mortucodld, mortdiab, morthipfx, morthypr, mortcms, mortndi)) -> dat_ipums

#####manage nchs mort dat
dat_nchs <- fread(here('mort_dat_1986_2018_2019.csv'))

#make merging id
dat_nchs %>%
  mutate(nhispid=as.numeric(ifelse(year<1997, paste(19, publicid, sep=""), paste(year, substring(publicid, 5), sep="")))) -> dat_nchs

#####merge
merged <- merge(dat_ipums, dat_nchs, by=c("nhispid", "year"), all=T)

#####exclusions
#exclude those outside age ranges (age has 0.09% missing)
#exclude those ineligible for mortality follow-up and those <25 or 65+ and those who aren't sample adults from 1997-on (occ1995 and classwk not available for others after that time); #note, from 2015-2018, only sample adult and sample child participants were eligible for mortality follow-up
#exclude 1997-2000, when there's no data on whether business is incorporated or not
#we'll exclude 1992 Hispanic oversample (only 3517 additional respondents) (per advice of IPUMS) separately, as we want to include them in sensitivity analyses
merged %>%
  filter((age>=25 & age<65) & eligstat==1 & (is.na(astatflg) | astatflg==1) & (year<1997 | (year>2000))) -> dat_sub

#make variables
dat_sub %>%
  mutate_at(.vars=(vars(-c(ucod_leading, wgt_new, sa_wgt_new, mortwt, mortwtsa, psu, strata, nhispid, nhishid))),
            ~ifelse(. %in% c(91, 96, 97, 98, 99, 970, 980, 990, 997, 999, 7777, 8888, 9998, 9999), NA, .)) %>%
  mutate(classwk2=ifelse(classwk2 %in% (7:9), NA, classwk2), #these are real values in some other variables so we'll reset them here
         jobsecincorp=ifelse(jobsecincorp %in% (7:9), NA, jobsecincorp),
         racesr=ifelse(racesr==900, NA, racesr),
         lamtwrk=ifelse(lamtwrk %in% 7:9, NA, lamtwrk)) %>%
  mutate(mortwt_f=ifelse(year==1986, mortwt, #per NCHS advice, use basic weight in 1986 (IPUMS), person mort weights 1987-1996 (NCHS), then sample adult mort weights 1997-2018 (NCHS)
                         ifelse(year>1986 & year<=1996, wgt_new, sa_wgt_new)), #https://www.cdc.gov/nchs/data/datalinkage/2019NDI-Linkage-Methods-and-Analytic-Considerations-508.pdf
         hispanic_oversample=ifelse((year==1992 & substr(nhispid, 1, 4)==1991), 1, 0),
         int_year_fin=ifelse(year<2001, year, #1986-1996, separate interview year unavailable
                             ifelse(year>=2001 & !is.na(intervwyr), intervwyr, year)), #if interview year is missing (<0.1%), assume it happened in survey-wave year (true in >99% of cases)
         dead=ifelse(mortstat==1, 1, 0),
         time=ifelse(dead==0, 2020 - int_year_fin, dodyear + 1 - int_year_fin), #deaths thru end of 2019, ie, 2020; assume deaths occurred at end of each year (1986 deaths occurred at end of 1986, ie, 1987)
         dead_86_96_05=ifelse(int_year_fin>1996, NA, #time and death indicators for 1986-1996 panel
                              ifelse(int_year_fin<=1996 & (is.na(dodyear) | dodyear>2004), 0,
                                    ifelse(int_year_fin<=1996 & dodyear<=2004, 1, NA))),
         time_86_96_05=ifelse(int_year_fin>1996, NA,
                              ifelse(dead_86_96_05==1, dodyear + 1 - int_year_fin,
                                     ifelse(dead_86_96_05==0, 2005-int_year_fin, NA))),
         race=ifelse(racesr==100, "White", 
                     ifelse(racesr==200, "Black", "Other")),
         hispeth_bin=ifelse(hispeth!=10, 1, 0),
         race_h=factor(ifelse(hispeth!=10, "Hispanic",
                              ifelse(hispeth==10 & race=="White", "NH white",
                                     ifelse(hispeth==10 & race=="Black", "NH black", "NH other"))), levels=c("NH white", "NH black", "Hispanic", "NH other")),
         poc=ifelse(race_h=="NH white", "NH white", "POC"),
         sex=ifelse(sex==1, "male", "female"),
         race_h_sex=factor(ifelse(race_h=="NH white" & sex=="male", "NH white men",
                                  ifelse(race_h=="NH white" & sex=="female", "NH white women",
                                         ifelse(race_h=="NH black" & sex=="male", "NH black men",
                                                ifelse(race_h=="NH black" & sex=="female", "NH black women",
                                                       ifelse(race_h=="Hispanic" & sex=="male", "Hispanic men",
                                                              ifelse(race_h=="Hispanic" & sex=="female", "Hispanic women",
                                                                     ifelse(race_h=="NH other" & sex=="male", "NH other men", "NH other women"))))))),
                           levels=c("NH white men", "NH white women", "NH black men", "NH black women", "Hispanic men", "Hispanic women", "NH other men", "NH other women")),
         educ=factor(ifelse(educrec1<13, "<HS",
                            ifelse(educrec1==13, "HS",
                                   ifelse(educrec1==14, "Some college", "College+"))), levels=c("<HS", "HS", "Some college", "College+")),
         marital_tri=factor(ifelse(marstat %in% 10:12, "Married",
                                   ifelse(marstat %in% 20:40, "Widowed/divorced/separated", 
                                          ifelse(marstat==50, "Single", NA))), levels=c("Married", "Single", "Widowed/divorced/separated")),
         region=ifelse(region==1, "NE", 
                       ifelse(region==2, "MW",
                              ifelse(region==3, "S", "W"))),
         incimp1_rev=ifelse(year<2007, NA,
                            ifelse(incimp1>66, 66, incimp1)),
         managers=ifelse(occ1995>=102 & occ1995<=104, 1, 0),
         class=factor(ifelse(empstat==220, "NLFs",
                             ifelse((empstat>=200 & empstat<=214) | (year<2001 & classwk>=20 & classwk<=34 & managers!=1) | (year>=2001 & classwk2>=1 & classwk2<=4 & managers!=1), "Wrks",
                                    ifelse((year<2001 & classwk>=20 & classwk<=34 & managers==1) | (year>=2001 & classwk2>=1 & classwk2<=4 & managers==1), "Mans",
                                           ifelse((year<2001 & classwk==41) | (year>=2001 & classwk2==5 & jobsecincorp==2), "IBOs",
                                                  ifelse((year<2001 & (classwk==42 | classwk==50)) | ((year>=2001 & classwk2==6) | (year>=2001 & classwk2==5 & jobsecincorp==1)), "UBOs", NA))))), 
                     levels=c("IBOs", "UBOs", "Mans", "Wrks", "NLFs")),
         class_gender=factor(ifelse(class=="Wrks" & sex=="male", "Male wrks",
                                    ifelse(class=="Wrks" & sex=="female", "Female wrks",
                                           ifelse(class=="Mans" & sex=="male", "Male mans",
                                                  ifelse(class=="Mans" & sex=="female", "Female mans",
                                                         ifelse(class=="UBOs" & sex=="male", "Male UBOs",
                                                                ifelse(class=="UBOs" & sex=="female", "Female UBOs",
                                                                       ifelse(class=="IBOs" & sex=="male", "Male IBOs",
                                                                              ifelse(class=="IBOs" & sex=="female", "Female IBOs",
                                                                                     ifelse(class=="NLFs" & sex=="male", "Male NLFs",
                                                                                            ifelse(class=="NLFs" & sex=="female", "Female NLFs", NA)))))))))),
                             levels=c("Male IBOs", "Male UBOs", "Male mans", "Male wrks", "Male NLFs",
                                      "Female IBOs", "Female UBOs", "Female mans", "Female wrks", "Female NLFs")),
         class_poc=factor(ifelse(class=="Wrks" & poc=="NH white", "NH white wrks",
                                 ifelse(class=="Wrks" & poc=="POC", "POC wrks",
                                        ifelse(class=="Mans" & poc=="NH white", "NH white mans",
                                               ifelse(class=="Mans" & poc=="POC", "POC mans",
                                                      ifelse(class=="UBOs" & poc=="NH white", "NH white UBOs",
                                                             ifelse(class=="UBOs" & poc=="POC", "POC UBOs",
                                                                    ifelse(class=="IBOs" & poc=="NH white", "NH white IBOs",
                                                                           ifelse(class=="IBOs" & poc=="POC", "POC IBOs", 
                                                                                  ifelse(class=="NLFs" & poc=="NH white", "NH white NLFs", 
                                                                                         ifelse(class=="NLFs" & poc=="POC", "POC NLFs", NA)))))))))),
                          levels=c("NH white IBOs", "NH white UBOs", "NH white mans", "NH white wrks", "NH white NLFs",
                                   "POC IBOs", "POC UBOs", "POC mans", "POC wrks", "POC NLFs")),
         class_educ=factor(ifelse(class=="Wrks" & (educ=="<HS" | educ=="HS"), "\u2264HS wrks",
                                 ifelse(class=="Wrks" & (educ=="Some college" | educ=="College+"), ">HS wrks",
                                        ifelse(class=="Mans" & (educ=="<HS" | educ=="HS"), "\u2264HS mans",
                                               ifelse(class=="Mans" & (educ=="Some college" | educ=="College+"), ">HS mans",
                                                      ifelse(class=="UBOs" & (educ=="<HS" | educ=="HS"), "\u2264HS UBOs",
                                                             ifelse(class=="UBOs" & (educ=="Some college" | educ=="College+"), ">HS UBOs",
                                                                    ifelse(class=="IBOs" & (educ=="<HS" | educ=="HS"), "\u2264HS IBOs",
                                                                           ifelse(class=="IBOs" & (educ=="Some college" | educ=="College+"), ">HS IBOs", 
                                                                                  ifelse(class=="NLFs" & (educ=="<HS" | educ=="HS"), "\u2264HS NLFs", 
                                                                                         ifelse(class=="NLFs" & (educ=="Some college" | educ=="College+"), ">HS NLFs", NA)))))))))),
                          levels=c(">HS IBOs", ">HS UBOs", ">HS mans", ">HS wrks", ">HS NLFs",
                                   "\u2264HS IBOs", "\u2264HS UBOs", "\u2264HS mans", "\u2264HS wrks", "\u2264HS NLFs")),
         class_sens=factor(ifelse(year<2001 | year>2003, NA, #identical to class variable
                                  ifelse(empstat==220, "NLFs",
                                         ifelse((empstat>=200 & empstat<=214) | (classwk>=20 & classwk<=34 & managers!=1), "Wrks",
                                                ifelse((classwk>=20 & classwk<=34 & managers==1), "Mans",
                                                       ifelse(classwk==40 & jobsecincorp==2, "IBOs",
                                                              ifelse((classwk==40 & jobsecincorp==1) | classwk==50, "UBOs", NA)))))), 
                          levels=c("IBOs", "UBOs", "Mans", "Wrks", "NLFs")),
         class_occ=factor(ifelse(empstat==220, "NLFs",
                                 ifelse(empstat>=200 & empstat<=214, "Unemps",
                                        ifelse((year<2001 & classwk>=20 & classwk<=34 & occ1995>=900) | (year>=2001 & classwk2>=1 & classwk2<=4 & occ1995>=900), "Blue collar/farm",
                                               ifelse((year<2001 & classwk>=20 & classwk<=34 & occ1995>=600 & occ1995<900) | (year>=2001 & classwk2>=1 & classwk2<=4 & occ1995>=600 & occ1995<900), "Services",
                                                      ifelse((year<2001 & classwk>=20 & classwk<=34 & occ1995>=200 & occ1995<600) | (year>=2001 & classwk2>=1 & classwk2<=4 & occ1995>=200 & occ1995<600), "White collar",
                                                             ifelse((year<2001 & classwk>=20 & classwk<=34 & managers==1) | (year>=2001 & classwk2>=1 & classwk2<=4 & managers==1), "Mans",
                                                                    ifelse((year<2001 & classwk==41) | (year>=2001 & classwk2==5 & jobsecincorp==2), "IBOs",
                                                                           ifelse((year<2001 & (classwk==42 | classwk==50)) | ((year>=2001 & classwk2==6) | (year>=2001 & classwk2==5 & jobsecincorp==1)), "UBOs", NA)))))))), 
                     levels=c("IBOs", "UBOs", "Mans", "White collar", "Services", "Blue collar/farm", "Unemps", "NLFs"))) -> dat_sub

#exclude 1 respondent who supposedly died the year before being interviewed (possibly due to NCHS data peturbation date of death: https://www.cdc.gov/nchs/data/datalinkage/public-use-2015-linked-mortality-file-description.pdf)
dat_sub %>%
  filter(time>=0) -> dat_sub

#make dataset without hispanic oversample for main analyses
dat_sub %>%
  filter(hispanic_oversample==0) -> dat_sub_no_hisp
```

# Missingness 

Age missingness is in full dataset among years and samples of interest; missingness for other variables is among those ages 25-64

```{r message=FALSE, warning=FALSE}
#age in full dataset
merged %>%
  filter(eligstat==1 & (is.na(astatflg) | astatflg==1) & (year<1997 | (year>2000)) & 
           !(year==1992 & substr(nhispid, 1, 4)==1991)) %>%
  mutate_at(.vars=(vars(-c(ucod_leading, mortwt, mortwtsa, psu, strata, nhispid, nhishid))), ~ifelse(. %in% c(91, 96, 97, 98, 99, 970, 980, 990, 7777, 8888, 9999), NA, .)) %>%
  dplyr::select('age') %>%
  miss_var_summary() -> dat_age_na

#other vars in subsetted dataset
dat_sub_no_hisp %>%
  dplyr::select(c('class', 'sex', 'race_h', 'educ', 'marital_tri', 'region')) %>%
  miss_var_summary() -> dat_other_na

#print
kable(bind_rows(dat_age_na, dat_other_na), digits=2, col.names=c("Variable", "Number missing", "Percent missing")) %>%
  kable_styling('striped')

#create complete-case dataset for rest of analyses
dat_sub_no_hisp <- dat_sub_no_hisp[complete.cases(dat_sub_no_hisp[,c('age', 'time', 'class', 'sex', 'race_h', 'educ', 'marital_tri', 'region')]),]
```

# Descriptives

Weighted and excluding missingness unless otherwise noted

## Unweighted and stratified table one 

```{r message=FALSE, warning=FALSE}
#vars of interest
vars <- c('sex', 'race_h', 'educ', 'marital_tri', 'region', 'age', 'year', 'time', 'dead')
catvars <- c("sex", 'dead')
nonorm <- c('age', 'year', 'time')

x <- CreateTableOne(data = dat_sub_no_hisp, vars=vars, factorVars=catvars, strata='class', includeNA=TRUE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, test=FALSE)
kable(x) %>%
  kable_styling(c("striped", "condensed"))
```

## Stratified table one 

```{r message=FALSE, warning=FALSE}
dat_sub_no_hisp_svy <- svydesign(ids = ~ psu,
                         strata = ~ strata, 
                         weights = ~ mortwt_f,
                         nest=TRUE, 
                         data=dat_sub_no_hisp)

x <- svyCreateTableOne(data = dat_sub_no_hisp_svy, vars=vars, factorVars=catvars, strata='class', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
kable(x) %>%
  kable_styling(c("striped", "condensed"))
```

## Stratified class composition of gender-races 

### Table

```{r message=FALSE, warning=FALSE}
x <- svyCreateTableOne(data=dat_sub_no_hisp_svy, vars='class', factorVars='class', strata='race_h_sex', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
kable(x[2:7,]) %>%
  kable_styling(c("striped", "condensed"))
```

### Mosaic plot

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
ggplot(data=dat_sub_no_hisp) +
  geom_mosaic(aes(x=product(class, race_h_sex), fill=class, weight=mortwt_f)) +
  ylab("") +
  xlab("")  +
  scale_y_productlist(expand=c(0,0.01)) +
  scale_x_productlist(expand=c(0,0.01)) +
  scale_fill_manual(values=brewer.pal(n = 12, name = "Paired")[c(2,8,4,10,6)]) +
  theme_classic() +
  theme(axis.line=element_blank(), axis.title=element_blank(), legend.position="none",
        axis.text.y=element_text(size=10, color="black"), axis.text.x=element_text(size=10, color="black", angle=25, vjust=1, hjust=1)) 

ggsave('mosaic.png', dpi=600, height=5, width=10)
```


## Proportions in classes over time 

Using interview year rather than survey-wave year. Excluded 2019 because <50 respondents interviewed in 2019.

```{r message=FALSE, warning=FALSE}
dat_sub_no_hisp %>%
  filter(int_year_fin!=2019) %>%
  group_by(class, int_year_fin) %>%
  summarise(n = sum(mortwt_f)) %>%
  ungroup() %>%
  group_by(int_year_fin) %>%
  mutate(prop = n / sum(n)) -> propped

ggplot(propped, aes(x=int_year_fin, y=prop, group=class, color=class, label=class)) +
  geom_dl(method='last.qp') +
  geom_line() +
  scale_color_manual(values=brewer.pal(n = 12, name = "Paired")[c(2,8,4,10,6)]) +
  scale_x_continuous(breaks=c(1986, 1994, 2002, 2010, 2018), expand=expansion(mult=c(0,0.08))) +
  scale_y_continuous(limits=c(0, 0.63), expand=expansion(mult=c(0,0))) +
  xlab("Interview year") +
  ylab("Proportion") +
  theme_light() +
  theme(legend.position = "none")

ggsave('proportions_over_time.png', dpi=600)
```

## Family income distribution across classes from 2007-2018 

Income coding inconsistent prior to 2007. Unadjusted for inflation. Row label shows lower bound of category.

```{r message=FALSE, warning=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, year>=2007), vars='incimp1_rev', factorVars='incimp1_rev', strata='class', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
row.names(x)[3:23] <- c(paste0(seq(0,95,5), "k"),">=100k")
kable(x[3:23,]) %>%
  kable_styling(c("striped", "condensed"))
```

## Number of employees at work across classes from 2001-2018

Excluding those not working. IBOs 4x as likely as UBOs to be in firms with 10+ employees. It probably makes sense that managers and workers are much more likely than business owners to work in larger firms, since any given firm, regardless of size, will only have one owner (in theory), whereas such firms will have multiple workers. For example, in a society with 5 firms, 4 of which have 5 employees and 1 of which has 1000 employees, 80% of owners will come from firms with 5 employees, but 98% (1000/1020) of workers will come from firms with 1000+ employees.

```{r message=FALSE, warning=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, year>=2001 & numemps!=0 & class!="NLFs"), vars='numemps', factorVars='numemps', strata='class', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
row.names(x)[3:10] <- c("1-9", "10-24", "25-49", "50-99", "100-249", "250-499", "500-999", "1000+")
kable(x[3:10,1:4]) %>%
  kable_styling(c("striped", "condensed"))
```

## Reason not in labor force

### Main reason for not working last week

Variable only available from 2004-on.

```{r warning=FALSE, message=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, year>=2004 & empstat==220 & whynowk2!=0), vars='whynowk2', factorVars='whynowk2', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
row.names(x)[3:12] <- c("Keeping house", "Going to school", "Retired", "Unable to work for health reasons", "Disabled", "On layoff", "On planned vacation", "On family leave", "Have job/contract; off season", "Other")
kable(x[3:12,], digits=2, col.names="Percent") %>% 
  kable_styling("striped")
```

### Limited in kind/amount of work can do

Variable only available from 1997-on.

```{r warning=FALSE, message=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, year>=1997), vars='lamtwrk', factorVars='lamtwrk', strata='empstat', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
row.names(x)[3:5] <- c("Not limited", "Limited", "Unable to work")
kable(x[3:5,], col.names=c("Working for pay", "Working w/o pay", "With job, but not at work", "Not employed", "NILF")) %>% 
  kable_styling("striped")
```

# Class and all-cause mortality

```{r}
#plotting function
plotted <- function(datted, 
                    classed_repped=34,
                    classed=c(rep("IBOs", classed_repped), rep("UBOs", classed_repped), rep("Mans", classed_repped), rep("Wrks", classed_repped), rep("NLFs", classed_repped)),
                    repped=5,
                    grpd, 
                    grps, 
                    dummied=c("IBOs", "UBOs", "Mans", "Wrks", "NLFs"),
                    cexed=0.85,
                    vjusted=1,
                    val=brewer.pal(n = 12, name = "Paired")[c(2,4,6,8,10)], 
                    bottom=0.50){
  
  dat_surv <- rbind(data.frame(time=datted$time,
                               surved=datted$surv,
                               se=datted$std.err,
                               class=classed,
                               grp=rep(grpd, grps)),
                    data.frame(time=rep(0, repped), surved=rep(1, repped), se=rep(0, repped), class=dummied, grp=rep(grpd, grps)))
  
  ggplot(dat_surv, aes(x=time, y=surved, group=class, color=class, fill=class, label=class)) + 
    facet_wrap(~grp) +
    geom_line() +
    geom_ribbon(aes(ymin=surved-1.96*se, ymax=surved+1.96*se, color=class), alpha=0.2, lty=0) +
    geom_dl(method=list('last.bumpup', cex=cexed, vjust=vjusted)) +
    scale_color_manual(values=val) +
    scale_fill_manual(values=val) +
    xlab("Follow-up time (years)") +
    ylab("Probability of survival") +
    scale_x_continuous(limits=c(0, 34), breaks=seq(1, 34, 3), expand=expansion(mult=c(0,0.11))) +
    scale_y_continuous(limits=c(bottom, 1.0), expand=expansion(mult=c(0,0))) +
    theme_light() +
    theme(legend.position="none", strip.background=element_rect(color="darkgrey", fill=NA), 
          strip.text=element_text(color="black"), panel.grid.minor.x = element_blank())
}

#group plotting function
plotted_group <- function(datted, 
                    classed_repped=34,
                    classed=rep(c(rep("IBOs", classed_repped), rep("UBOs", classed_repped), rep("Mans", classed_repped), rep("Wrks", classed_repped), rep("NLFs", classed_repped)), 2),
                    repped=10,
                    grpd1, 
                    grps1=170, 
                    grpd2,
                    grps2=170,
                    dummied=c("IBOs", "UBOs", "Mans", "Wrks", "NLFs"),
                    levelled=c(grpd1, grpd2),
                    cexed=0.85,
                    vjusted=1,
                    val=brewer.pal(n = 12, name = "Paired")[c(2,4,6,8,10)], 
                    bottom=0.50){
  
  dat_surv <- rbind(data.frame(time=rep(0, repped),
                               surved=rep(1, repped), 
                               se=rep(0, repped), 
                               class=dummied,  
                               grp=c(rep(grpd1, 5), rep(grpd2, 5))),
                    data.frame(time=datted$time,
                               surved=datted$surv,
                               se=datted$std.err,
                               class=classed,
                               grp=c(rep(grpd1, grps1), rep(grpd2, grps2)))) %>%
    mutate(grp=factor(grp, levels=levelled))
  
  ggplot(dat_surv, aes(x=time, y=surved, group=class, color=class, fill=class, label=class)) + 
    facet_wrap(~grp) +
    geom_line() +
    geom_ribbon(aes(ymin=surved-1.96*se, ymax=surved+1.96*se, color=class), alpha=0.2, lty=0) +
    geom_dl(method=list('last.bumpup', cex=cexed, vjust=vjusted)) +
    scale_color_manual(values=val) +
    scale_fill_manual(values=val) +
    xlab("Follow-up time (years)") +
    ylab("Probability of survival") +
    scale_x_continuous(limits=c(0, 34), breaks=seq(1, 34, 3), expand=expansion(mult=c(0,0.11))) +
    scale_y_continuous(limits=c(bottom, 1.0), expand=expansion(mult=c(0,0))) +
    theme_light() +
    theme(legend.position="none", strip.background=element_rect(color="darkgrey", fill=NA), 
          strip.text=element_text(color="black"), panel.grid.minor.x = element_blank())
}

#survdiff function
survdiffed <- function(datted, 
                       timed=34){
  
  #end of f/u
  point <- summary(datted, times=timed)
  
  #pb-cap
  pb <- 100*(point$surv[2] - point$surv[1])
  pb_lower <- 100*(pb/100 - 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  pb_upper <- 100*(pb/100 + 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
    
  #mc-cap
  mc <- 100*(point$surv[3] - point$surv[1])
  mc_lower <- 100*(mc/100 - 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  mc_upper <- 100*(mc/100 + 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  
  #worker-cap
  wc <- 100*(point$surv[4] - point$surv[1])
  wc_lower <- 100*(wc/100 - 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  wc_upper <- 100*(wc/100 + 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  
  #nilf-cap
  nc <- 100*(point$surv[5] - point$surv[1])
  nc_lower <- 100*(nc/100 - 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  nc_upper <- 100*(nc/100 + 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  
  rbind(c(pb, pb_lower, pb_upper), c(mc, mc_lower, mc_upper), c(wc, wc_lower, wc_upper), c(nc, nc_lower, nc_upper))
  
}

#survdiff function for groups
survdiffed_group <- function(datted,
                             timed=34){
  
  #end of f/u
  point <- summary(datted, times=timed)
  
  #pb-cap
  pb <- 100*(point$surv[2] - point$surv[1])
  pb_lower <- 100*(pb/100 - 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  pb_upper <- 100*(pb/100 + 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  
  #mc-cap
  mc <- 100*(point$surv[3] - point$surv[1])
  mc_lower <- 100*(mc/100 - 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  mc_upper <- 100*(mc/100 + 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  
  #worker-cap
  wc <- 100*(point$surv[4] - point$surv[1])
  wc_lower <- 100*(wc/100 - 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  wc_upper <- 100*(wc/100 + 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  
  #nilf-cap
  nc <- 100*(point$surv[5] - point$surv[1])
  nc_lower <- 100*(nc/100 - 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  nc_upper <- 100*(nc/100 + 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  
  #cap2-cap
  cap2 <- 100*(point$surv[6] - point$surv[1])
  cap2_lower <- 100*(cap2/100 - 1.96*sqrt(point$std.err[6]^2 + point$std.err[1]^2))
  cap2_upper <- 100*(cap2/100 + 1.96*sqrt(point$std.err[6]^2 + point$std.err[1]^2))
  
  #pb2-cap
  pb2 <- 100*(point$surv[7] - point$surv[1])
  pb2_lower <- 100*(pb2/100 - 1.96*sqrt(point$std.err[7]^2 + point$std.err[1]^2))
  pb2_upper <- 100*(pb2/100 + 1.96*sqrt(point$std.err[7]^2 + point$std.err[1]^2))
  
  #man2-cap
  man2 <- 100*(point$surv[8] - point$surv[1])
  man2_lower <- 100*(man2/100 - 1.96*sqrt(point$std.err[8]^2 + point$std.err[1]^2))
  man2_upper <- 100*(man2/100 + 1.96*sqrt(point$std.err[8]^2 + point$std.err[1]^2))
  
  #wrk2-cap
  wrk2 <- 100*(point$surv[9] - point$surv[1])
  wrk2_lower <- 100*(wrk2/100 - 1.96*sqrt(point$std.err[9]^2 + point$std.err[1]^2))
  wrk2_upper <- 100*(wrk2/100 + 1.96*sqrt(point$std.err[9]^2 + point$std.err[1]^2))
  
  #nilf2-cap
  nilf2 <- 100*(point$surv[10] - point$surv[1])
  nilf2_lower <- 100*(nilf2/100 - 1.96*sqrt(point$std.err[10]^2 + point$std.err[1]^2))
  nilf2_upper <- 100*(nilf2/100 + 1.96*sqrt(point$std.err[10]^2 + point$std.err[1]^2))
  
  rbind(c(pb, pb_lower, pb_upper), c(mc, mc_lower, mc_upper), c(wc, wc_lower, wc_upper), c(nc, nc_lower, nc_upper),
        c(cap2, cap2_lower, cap2_upper), c(pb2, pb2_lower, pb2_upper), c(man2, man2_lower, man2_upper), c(wrk2, wrk2_lower, wrk2_upper), c(nilf2, nilf2_lower, nilf2_upper))
  
}

#survdiff function for worker subdivision
survdiffed_sub <- function(datted, 
                           timed=34){
  
  #end of f/u
  point <- summary(datted, times=timed)
  
  #pb-cap
  pb <- 100*(point$surv[2] - point$surv[1])
  pb_lower <- 100*(pb/100 - 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  pb_upper <- 100*(pb/100 + 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  
  #mc-cap
  mc <- 100*(point$surv[3] - point$surv[1])
  mc_lower <- 100*(mc/100 - 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  mc_upper <- 100*(mc/100 + 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  
  #whitecollar-cap
  wc <- 100*(point$surv[4] - point$surv[1])
  wc_lower <- 100*(wc/100 - 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  wc_upper <- 100*(wc/100 + 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  
  #services-cap
  servs <- 100*(point$surv[5] - point$surv[1])
  servs_lower <- 100*(servs/100 - 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  servs_upper <- 100*(servs/100 + 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  
  #bluecollar-cap
  bc <- 100*(point$surv[6] - point$surv[1])
  bc_lower <- 100*(bc/100 - 1.96*sqrt(point$std.err[6]^2 + point$std.err[1]^2))
  bc_upper <- 100*(bc/100 + 1.96*sqrt(point$std.err[6]^2 + point$std.err[1]^2))
  
  #pb2-cap
  unemp <- 100*(point$surv[7] - point$surv[1])
  unemp_lower <- 100*(unemp/100 - 1.96*sqrt(point$std.err[7]^2 + point$std.err[1]^2))
  unemp_upper <- 100*(unemp/100 + 1.96*sqrt(point$std.err[7]^2 + point$std.err[1]^2))
  
  #man2-cap
  nilf <- 100*(point$surv[8] - point$surv[1])
  nilf_lower <- 100*(nilf/100 - 1.96*sqrt(point$std.err[8]^2 + point$std.err[1]^2))
  nilf_upper <- 100*(nilf/100 + 1.96*sqrt(point$std.err[8]^2 + point$std.err[1]^2))
  
  rbind(c(pb, pb_lower, pb_upper), c(mc, mc_lower, mc_upper), c(wc, wc_lower, wc_upper), c(servs, servs_lower, servs_upper),
        c(bc, bc_lower, bc_upper), c(unemp, unemp_lower, unemp_upper), c(nilf, nilf_lower, nilf_upper))
  
}

#tidying regression output function
tidy_n <- function(modded=mod, 
                   rows=1:4, 
                   cols=c(1,2,7,8),
                   bind=binded,
                   nad=3, 
                   captioned="Ref: IBOs") {
  tidydf <- cbind(tidy(modded, exponentiate=T, conf.int=T)[rows, cols], bind)
  tidydf$N <- c(length(modded$residuals), rep(NA, nad))
  kable(tidydf, digits=c(2,2,2,2,1,1,1,1), caption=captioned, col.names=c(" ", "HR", "Lower", "Upper", "SD", "Lower", "Upper", "N")) %>%
    kable_styling("striped") %>%
    add_header_above(c(" ", "HR"=3, "SD per 100 at end of f/u"=3, " "))
}
```

## Overall and subdivided by occupation

### Less-adjusted (age, gender, year)

#### Distribution of IPW

##### Overall

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  mutate(sw_over=ipwpoint(exposure=class,
                          family="multinomial",
                          link="logit",
                          numerator=~1,
                          denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 5), 
                          data = dat_sub_no_hisp,
                          weights=mortwt_f,
                          trace=FALSE,
                          maxit=1000)$ipw.weights,
         sw_f_over=sw_over*mortwt_f)  -> dat_sub_no_hisp

summary(dat_sub_no_hisp$sw_over)
```

##### Subdivided

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  mutate(sw_sub=ipwpoint(exposure=class_occ,
                         family="multinomial",
                         link="logit",
                         numerator=~1,
                         denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 5), 
                         data = dat_sub_no_hisp,
                         weights=mortwt_f,
                         maxit=1000,
                         trace=FALSE)$ipw.weights,
         sw_f_sub=sw_sub*mortwt_f)  -> dat_sub_no_hisp

summary(dat_sub_no_hisp$sw_sub)
```

#### Covariate balance after weighting

##### Overall

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=9}
covs <- subset(dat_sub_no_hisp, select=c('age', 'year', 'sex'))
               
love.plot(covs,
          treat=dat_sub_no_hisp$class, 
          weights=dat_sub_no_hisp$sw_over, 
          threshold=c(m=0.1), 
          binary='std',
          which=.all) +
  facet_wrap(~treat, nrow=2) +
  scale_color_manual(values=c("#77AADD", "#114477")) +
  theme_light() +
  theme(legend.title=element_blank(), plot.title=element_blank(), 
        strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black")) +
  scale_y_discrete(labels=rev(c("Age", "Year", "Gender"))) 
```

##### Subdivided 

```{r warning=FALSE, message=FALSE, fig.height=9, fig.width=14}
covs <- subset(dat_sub_no_hisp, select=c('age', 'year', 'sex'))
               
love.plot(covs,
          treat=dat_sub_no_hisp$class_occ, 
          weights=dat_sub_no_hisp$sw_sub, 
          threshold=c(m=0.1), 
          binary='std',
          which=.all) +
  facet_wrap(~treat, nrow=4) +
  scale_color_manual(values=c("#77AADD", "#114477")) +
  theme_light() +
  theme(legend.title=element_blank(), plot.title=element_blank(), 
        strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black")) +
  scale_y_discrete(labels=rev(c("Age", "Year", "Gender"))) 
```

#### Survival 

```{r warning=FALSE, message=FALSE}
#overall
kapped_over <- survfit(Surv(time,dead)~class, robust=T, w=sw_f_over, data=dat_sub_no_hisp, se=T)
mod_over <- coxph(Surv(time,dead)~class, robust=T, w=sw_f_over, data=dat_sub_no_hisp)
binded_over <- survdiffed(datted=kapped_over)

#subdivided workers
kapped_sub <- survfit(Surv(time,dead)~class_occ, robust=T, w=sw_f_sub, data=dat_sub_no_hisp, se=T)
mod_sub <- coxph(Surv(time,dead)~class_occ, robust=T, w=sw_f_sub, data=dat_sub_no_hisp)
binded_sub <- survdiffed_sub(datted=kapped_sub)
```

##### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.75}
plotted(datted=kapped_over, 
        grpd="Workers undivided", 
        grps=170) + 
  scale_x_continuous(limits=c(0, 34), breaks=seq(1, 34, 3), expand=expansion(mult=c(0,0.16))) +
plotted(datted=kapped_sub, 
        classed=factor(c(rep("IBOs", 34), rep("UBOs", 34), rep("Mans", 34), rep("WC", 34), rep("Servs", 34), 
                         rep("BC/farm", 34), rep("Unemps", 34), rep("NLFs", 34)), 
                       levels=c("IBOs", "UBOs", "Mans", "WC", "BC/farm", "Servs", "Unemps", "NLFs")),
        grpd="Workers subdivided",
        grps=272, 
        repped=8,
        dummied=c("IBOs", "UBOs", "Mans", "WC", "Servs", "BC/farm", "Unemps", "NLFs"),
        val=c(brewer.pal(n = 12, name = "Paired")[c(2,8,4)], brewer.pal(n=9, name="Purples")[c(5,6,7,8)], brewer.pal(n = 12, name = "Paired")[c(6)])) + 
  scale_x_continuous(limits=c(0, 34), breaks=seq(1, 34, 3), expand=expansion(mult=c(0,0.16))) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) 

ggsave("overall_subdivided.png", dpi=600, height=4, width=8.75)
```

##### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_over, bind=binded_over)
tidy_n(modded=mod_sub, bind=binded_sub, rows=1:7, nad=6)
```

### More-adjusted (age, gender, year, education, marital status, region, and racialized group)

#### Distribution of IPW

Truncated at 0.5% and 99.5%.

##### Overall

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  mutate(sw_over_adj=ipwpoint(exposure=class,
                              family="multinomial",
                              link="logit",
                              numerator=~1,
                              denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 5) + educ + marital_tri + region + race_h, 
                              data = dat_sub_no_hisp,
                              weights=mortwt_f,
                              trace=FALSE,
                              maxit=1000,
                              trunc=0.005)$weights.trunc,
         sw_f_over_adj=sw_over_adj*mortwt_f)  -> dat_sub_no_hisp

summary(dat_sub_no_hisp$sw_over_adj)
```

##### Subdivided

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  mutate(sw_sub_adj=ipwpoint(exposure=class_occ,
                         family="multinomial",
                         link="logit",
                         numerator=~1,
                         denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 5) + educ + marital_tri + region + race_h, 
                         data = dat_sub_no_hisp,
                         weights=mortwt_f,
                         trace=FALSE,
                         maxit=1000,
                         trunc=0.005)$weights.trunc,
         sw_f_sub_adj=sw_sub_adj*mortwt_f)  -> dat_sub_no_hisp

summary(dat_sub_no_hisp$sw_sub_adj)
```

#### Covariate balance after weighting

##### Overall

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=9.5}
covs <- subset(dat_sub_no_hisp, select=c('age', 'year', 'sex', 'race_h', 'educ', 'marital_tri', 'region'))
               
love.plot(covs,
          treat=dat_sub_no_hisp$class, 
          weights=dat_sub_no_hisp$sw_over_adj, 
          threshold=c(m=0.1), 
          binary='std',
          which=.all) +
  facet_wrap(~treat, nrow=2) +
  scale_color_manual(values=c("#77AADD", "#114477")) +
  theme_light() +
  theme(legend.title=element_blank(), plot.title=element_blank(), 
        strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black")) +
  scale_y_discrete(labels=rev(c("Age", "Year", "Gender", 'NH white', "NH black", "Hispanic", "NH other", 
                                "<HS", "HS", "Some college", "College+", "Married", "Single", "Wid/div/sep", "MW", "NE", "S", "W"))) 
```

##### Subdivided 

```{r warning=FALSE, message=FALSE, fig.height=14, fig.width=14}
covs <- subset(dat_sub_no_hisp, select=c('age', 'year', 'sex', 'race_h', 'educ', 'marital_tri', 'region'))
               
love.plot(covs,
          treat=dat_sub_no_hisp$class_occ, 
          weights=dat_sub_no_hisp$sw_sub_adj, 
          threshold=c(m=0.1), 
          binary='std',
          which=.all) +
  facet_wrap(~treat, nrow=4) +
  scale_color_manual(values=c("#77AADD", "#114477")) +
  theme_light() +
  theme(legend.title=element_blank(), plot.title=element_blank(), 
        strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black")) +
  scale_y_discrete(labels=rev(c("Age", "Year", "Gender", 'NH white', "NH black", "Hispanic", "NH other", 
                                "<HS", "HS", "Some college", "College+", "Married", "Single", "Wid/div/sep", "MW", "NE", "S", "W"))) 
```

#### Survival 

```{r warning=FALSE, message=FALSE}
#overall
kapped_over <- survfit(Surv(time,dead)~class, robust=T, w=sw_f_over_adj, data=dat_sub_no_hisp, se=T)
mod_over <- coxph(Surv(time,dead)~class, robust=T, w=sw_f_over_adj, data=dat_sub_no_hisp)
binded_over <- survdiffed(datted=kapped_over)

#subdivided workers
kapped_sub <- survfit(Surv(time,dead)~class_occ, robust=T, w=sw_f_sub_adj, data=dat_sub_no_hisp, se=T)
mod_sub <- coxph(Surv(time,dead)~class_occ, robust=T, w=sw_f_sub_adj, data=dat_sub_no_hisp)
binded_sub <- survdiffed_sub(datted=kapped_sub)
```

##### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.75}
plotted(datted=kapped_over, 
        grpd="Workers undivided", 
        grps=170) + 
  scale_x_continuous(limits=c(0, 34), breaks=seq(1, 34, 3), expand=expansion(mult=c(0,0.16))) +
plotted(datted=kapped_sub, 
        classed=factor(c(rep("IBOs", 34), rep("UBOs", 34), rep("Mans", 34), rep("WC", 34), rep("Servs", 34), 
                         rep("BC/farm", 34), rep("Unemps", 34), rep("NLFs", 34)), 
                       levels=c("IBOs", "UBOs", "Mans", "WC", "Servs", "BC/farm", "Unemps", "NLFs")),
        grpd="Workers subdivided",
        grps=272, 
        repped=8,
        dummied=c("IBOs", "UBOs", "Mans", "WC", "Servs", "BC/farm", "Unemps", "NLFs"),
        vjusted=1.5,
        val=c(brewer.pal(n = 12, name = "Paired")[c(2,8,4)], brewer.pal(n=9, name="Purples")[c(5,6,7,8)], brewer.pal(n = 12, name = "Paired")[c(6)])) + 
  scale_x_continuous(limits=c(0, 34), breaks=seq(1, 34, 3), expand=expansion(mult=c(0,0.16))) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) 

ggsave("overall_subdivided_adj.png", dpi=600, height=4, width=8.75)
```

##### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_over, bind=binded_over)
tidy_n(modded=mod_sub, bind=binded_sub, rows=1:7, nad=6)
```

## Class-by-year stratification - 1986-1996 interview years w/ f/u through end of 2004 vs 2001-2019 interview years w/ f/u through end of 2019

### Less-adjusted

#### Distribution of IPW

##### 1986-1996

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(int_year_fin<=1996) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 3), 
                     data = subset(dat_sub_no_hisp, int_year_fin<=1996),
                     weights=mortwt_f,
                     maxit=1000,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_1986

summary(dat_sub_no_hisp_1986$sw)
```

##### 2001-2019

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(int_year_fin>=2001) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 3), 
                     data = subset(dat_sub_no_hisp, int_year_fin>=2001),
                     weights=mortwt_f,
                     maxit=1000,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_2001

summary(dat_sub_no_hisp_2001$sw)
```

#### Survival 

```{r warning=FALSE, message=FALSE}
#early
kapped_early <- survfit(Surv(time_86_96_05, dead_86_96_05)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_1986, se=T) 
mod_early <- coxph(Surv(time_86_96_05, dead_86_96_05)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_1986)
binded_early <- survdiffed(kapped_early, timed=19)

#late
kapped_late <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_2001, se=T)
mod_late <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_2001)
binded_late <- survdiffed(kapped_late, timed=19)
```

##### IPW survival plots

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.75}
plotted(datted=kapped_early,
        classed_repped=19,
        grpd="1986-1996 with follow-up through 2004", 
        grps=95, 
        bottom=0.75) + 
  scale_x_continuous(limits=c(0, 19), breaks=seq(1, 19, 2), expand=expansion(mult=c(0,0.11)))  +
plotted(datted=kapped_late,
        classed_repped=19,
        grpd="2001-2019 with follow-up through 2019", 
        grps=95, 
        bottom=0.75) + 
  scale_x_continuous(limits=c(0, 19), breaks=seq(1, 19, 2), expand=expansion(mult=c(0,0.11))) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())
                     
ggsave("change_over_time.png", dpi=600, height=4, width=8.75)
```

##### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_early, bind=binded_early, captioned="Ref: IBOs (1986-1996 subset)")
tidy_n(modded=mod_late, bind=binded_late, captioned="Ref: IBOs (2001-2019 subset)")
```

### More-adjusted

#### Distribution of IPW

Truncated at 0.5% and 99.5%.

##### 1986-1996

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(int_year_fin<=1996) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 3) + educ + marital_tri + region + race_h, 
                     data = subset(dat_sub_no_hisp, int_year_fin<=1996),
                     weights=mortwt_f,
                     trace=FALSE,
                     maxit=1000,
                     trunc=0.005)$weights.trunc,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_1986

summary(dat_sub_no_hisp_1986$sw)
```

##### 2001-2019

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(int_year_fin>=2001) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 3) + educ + marital_tri + region + race_h, 
                     data = subset(dat_sub_no_hisp, int_year_fin>=2001),
                     weights=mortwt_f,
                     trace=FALSE,
                     maxit=1000,
                     trunc=0.005)$weights.trunc,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_2001

summary(dat_sub_no_hisp_2001$sw)
```

#### Survival

```{r warning=FALSE, message=FALSE}
#early
kapped_early <- survfit(Surv(time_86_96_05, dead_86_96_05)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_1986, se=T) 
mod_early <- coxph(Surv(time_86_96_05, dead_86_96_05)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_1986)
binded_early <- survdiffed(kapped_early, timed=19)

#late
kapped_late <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_2001, se=T)
mod_late <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_2001)
binded_late <- survdiffed(kapped_late, timed=19)
```

##### IPW survival plots

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.75}
plotted(datted=kapped_early,
        classed_repped=19,
        grpd="1986-1996 with follow-up through 2004", 
        grps=95, 
        vjusted=1.5,
        bottom=0.75) + 
  scale_x_continuous(limits=c(0, 19), breaks=seq(1, 19, 2), expand=expansion(mult=c(0,0.11))) +
plotted(datted=kapped_late,
        classed_repped=19,
        grpd="2001-2019 with follow-up through 2019", 
        grps=95, 
        bottom=0.75) + 
  scale_x_continuous(limits=c(0, 19), breaks=seq(1, 19, 2), expand=expansion(mult=c(0,0.11))) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())
                     
ggsave("change_over_time_adj.png", dpi=600, height=4, width=8.75)
```

##### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_early, bind=binded_early, captioned="Ref: IBOs (1986-1996 subset)")
tidy_n(modded=mod_late, bind=binded_late, captioned="Ref: IBOs (2001-2019 subset)")
```

## Class-by-gender and class-by-racialized-group interaction

### Distribution of IPW

#### Gender

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  mutate(sw_gend=ipwpoint(exposure=class_gender,
                          family="multinomial",
                          link="logit",
                          numerator=~1,
                          denominator=~rcs(age, 3) + rcs(int_year_fin, 5), 
                          data = dat_sub_no_hisp,
                          weights=mortwt_f,
                          maxit=1000,
                          trace=FALSE)$ipw.weights,
         sw_f_gend=sw_gend*mortwt_f,
         gender=ifelse(sex=="female", "Women", "Men"))  -> dat_sub_no_hisp

summary(dat_sub_no_hisp$sw_gend)
```

#### Racialized group

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  mutate(sw_race=ipwpoint(exposure=class_poc,
                          family="multinomial",
                          link="logit",
                          numerator=~1,
                          denominator=~rcs(age, 3) + rcs(int_year_fin, 5), 
                          data = dat_sub_no_hisp,
                          weights=mortwt_f,
                          maxit=1000,
                          trace=FALSE)$ipw.weights,
         sw_f_race=sw_race*mortwt_f)  -> dat_sub_no_hisp

summary(dat_sub_no_hisp$sw_race)
```

### Survival

```{r warning=FALSE, message=FALSE}
#gender
kapped_gend <- survfit(Surv(time,dead)~class_gender, robust=T, w=sw_f_gend, data=dat_sub_no_hisp, se=T)
mod_gend <- coxph(Surv(time,dead)~class_gender, robust=T, w=sw_f_gend, data=dat_sub_no_hisp)
binded_gend <- survdiffed_group(kapped_gend)

#racialized group
kapped_race <- survfit(Surv(time,dead)~class_poc, robust=T, w=sw_f_race, data=dat_sub_no_hisp, se=T)
mod_race <- coxph(Surv(time,dead)~class_poc, robust=T, w=sw_f_race, data=dat_sub_no_hisp)
binded_race <- survdiffed_group(kapped_race)
```

#### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=7.5, fig.width=8.75}
plotted_group(kapped_gend, grpd1="Men", grpd2="Women", bottom=0.40) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
plotted_group(kapped_race, grpd1="NH white", grpd2="NH Black, Hispanic, and NH other", bottom=0.40) +
  plot_layout(ncol=1)

ggsave("gender_race_survival.png", dpi=600, height=7.5, width=8.75)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_gend, bind=binded_gend, rows=1:9, nad=8, captioned="Ref: male IBOs")
tidy_n(modded=mod_race, bind=binded_race, rows=1:9, nad=8, captioned="Ref: NH white IBOs")
```

## Class-by-education interaction

### Distribution of IPW

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  mutate(sw_educ=ipwpoint(exposure=class_educ,
                          family="multinomial",
                          link="logit",
                          numerator=~1,
                          denominator=~rcs(age, 3) + rcs(int_year_fin, 5), 
                          data=dat_sub_no_hisp,
                          weights=mortwt_f,
                          maxit=1000,
                          trace=FALSE)$ipw.weights,
         sw_f_educ=sw_educ*mortwt_f)  -> dat_sub_no_hisp

summary(dat_sub_no_hisp$sw_educ)
```

### Survival

```{r warning=FALSE, message=FALSE}
kapped_educ <- survfit(Surv(time,dead)~class_educ, robust=T, w=sw_f_educ, data=dat_sub_no_hisp, se=T)
mod_educ <- coxph(Surv(time,dead)~class_educ, robust=T, w=sw_f_educ, data=dat_sub_no_hisp)
binded_educ <- survdiffed_group(kapped_educ)
```

#### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.75}
plotted_group(kapped_educ, grpd1=">HS", grpd2="\u2264HS", levelled=c("\u2264HS", ">HS"))

ggsave("educ_survival.png", dpi=600, height=4, width=8.75)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_educ, bind=binded_educ, rows=1:9, nad=8, captioned="Ref: >HS IBOs")
```

## Class-by-age stratification - 25-44 vs 45-64 years old

### Distribution of IPW

#### 25-44

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(age<45) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 5), 
                     data = subset(dat_sub_no_hisp, age<45),
                     weights=mortwt_f,
                     maxit=1000,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_young

summary(dat_sub_no_hisp_young$sw)
```

#### 45-64

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(age>=45) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 5), 
                     data = subset(dat_sub_no_hisp, age>=45),
                     weights=mortwt_f,
                     maxit=1000,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_old

summary(dat_sub_no_hisp_old$sw)
```

### Survival

```{r warning=FALSE, message=FALSE}
#young
kapped_young <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_young, se=T)
mod_young <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_young)
binded_young <- survdiffed(kapped_young)

#old
kapped_old <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_old, se=T)
mod_old <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_old)
binded_old <- survdiffed(kapped_old)
```

#### IPW survival plots

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.75}
plotted(datted=kapped_young, grpd="25-44", vjust=1.25, bottom=0.2) + 
plotted(datted=kapped_old, grpd="45-64", bottom=0.2) +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())

ggsave("age_survival.png", dpi=600, height=4, width=8.75)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_young, bind=binded_young, captioned="Ref: IBOs (25-44 subset)")
tidy_n(modded=mod_old, bind=binded_old, captioned="Ref: IBOs (45-64 subset)")
```

## Sensitivity analyses

### Regression adjustment instead of IPW for minimally-adjusted overall Cox analyses

Similar to primary but slightly larger effect sizes for workers and UBOs.

```{r warning=FALSE, message=FALSE}
mod_reg <- coxph(Surv(time,dead)~class + rcs(age, 3) + sex + rcs(int_year_fin, 5), robust=T, w=mortwt_f, data=dat_sub_no_hisp)

kable(cbind(tidy(mod_reg, exponentiate=T, conf.int=T)[1:4, c(1,2,7,8)], c(length(mod_reg$residuals), NA, NA, NA)), 
      col.names=c(" ", "HR", "Lower", "Upper", "N"), caption="Ref: IBOs", digits=2) %>%
  kable_styling("striped")
```

### Checking that SEs in primary analyses align with SEs from other approaches

#### Accounting for survey design in overall Cox analyses (adjusting for confounders with IPW)

Basically identical to primary. It makes no difference that we subsetted the dataset by age above rather than within the svycoxph function because age subpopulations appear in all the clusters

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp_svy <- svydesign(ids = ~ psu,
                                 strata = ~ strata, 
                                 weights = ~ sw_f_over,
                                 nest=TRUE, 
                                 data=dat_sub_no_hisp)

mod_surv <- svycoxph(Surv(time,dead)~class, design=dat_sub_no_hisp_svy)

kable(cbind(tidy(mod_surv, exponentiate=T, conf.int=T)[1:4, c(1,2,7,8)], c(length(mod_surv$residuals), NA, NA, NA)), 
      col.names=c(" ", "HR", "Lower", "Upper", "N"), caption="Ref: IBOs", digits=2) %>%
  kable_styling("striped")
```

#### Clustering SEs at household level rather than leaving them unclustered

Also basically identical to primary.

```{r warning=FALSE, message=FALSE}
kapped_house <- survfit(Surv(time,dead)~class + cluster(nhishid), w=sw_f_over, data=dat_sub_no_hisp, se=T)
mod_house <- coxph(Surv(time,dead)~class + cluster(nhishid), w=sw_f_over, data=dat_sub_no_hisp)
binded_house <- survdiffed(kapped_house)
```

```{r}
tidy_n(modded=mod_house, bind=binded_house)
```

### Including Hispanic oversample to examine effects on point estimates 

Basically identical to primary, as Hispanic oversample is small (<0.5% of full sample).

#### Distribution of IPW

```{r warning=FALSE, message=FALSE}
dat_sub %>%
  filter(!is.na(class) & !is.na(age)) %>%
  mutate(sw_hisp=ipwpoint(exposure=class,
                          family="multinomial",
                          link="logit",
                          numerator=~1,
                          denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 5), 
                          data = subset(dat_sub, !is.na(class) & !is.na(age)),
                          weights=mortwt_f,
                          maxit=1000,
                          trace=FALSE)$ipw.weights,
         sw_f_hisp=sw_hisp*mortwt_f)  -> dat_sub_hisp_overall

summary(dat_sub_hisp_overall$sw_hisp)
```

```{r warning=FALSE, message=FALSE}
kapped_hisp <- survfit(Surv(time,dead)~class, robust=T, w=sw_hisp, data=dat_sub_hisp_overall, se=T)
mod_hisp <- coxph(Surv(time,dead)~class, robust=T, w=sw_hisp, data=dat_sub_hisp_overall)
binded_hisp <- survdiffed(kapped_hisp)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_hisp, bind=binded_hisp)
```

### Class by year interaction with regression adjustment instead of period stratification and IPW

Similar to primary. Used 2018 instead of 2019 as end-of-follow-up year because <50 respondents were interviewed in 2019.

#### Less-adjusted

```{r warning=FALSE, message=FALSE}
mod_1986 <- coxph(Surv(time,dead)~class*I(int_year_fin - 1986) + rcs(age, 3) + sex, robust=T, w=mortwt_f, data=dat_sub_no_hisp)

kable(cbind(tidy(mod_1986, exponentiate=T, conf.int=T)[1:4, c(1,2,7,8)], 
            exp((rbind(confint(glht(mod_1986, linfct = c("classUBOs + 32*classUBOs:I(int_year_fin - 1986) = 0")))$confint[1:3],
                       confint(glht(mod_1986, linfct = c("classMans + 32*classMans:I(int_year_fin - 1986) = 0")))$confint[1:3],
                       confint(glht(mod_1986, linfct = c("classWrks + 32*classWrks:I(int_year_fin - 1986) = 0")))$confint[1:3],
                       confint(glht(mod_1986, linfct = c("classNLFs + 32*classNLFs:I(int_year_fin - 1986) = 0")))$confint[1:3]))),
            c(length(mod_1986$residuals), NA, NA, NA)),
      col.names=c(" ", "HR", "Lower", "Upper", "HR", "Lower", "Upper", "N"),
      caption="Ref: IBOs in given year",
      digits=2) %>%
  kable_styling("striped") %>%
  add_header_above(c(" ", "HR in 1986"=3, "HR in 2018"=3, " "))
```

#### More-adjusted

```{r warning=FALSE, message=FALSE}
mod_1986 <- coxph(Surv(time,dead)~class*I(int_year_fin - 1986) + rcs(age, 3) + sex + educ + marital_tri + region + race_h, robust=T, w=mortwt_f, data=dat_sub_no_hisp)

kable(cbind(tidy(mod_1986, exponentiate=T, conf.int=T)[1:4, c(1,2,7,8)], 
            exp((rbind(confint(glht(mod_1986, linfct = c("classUBOs + 32*classUBOs:I(int_year_fin - 1986) = 0")))$confint[1:3],
                       confint(glht(mod_1986, linfct = c("classMans + 32*classMans:I(int_year_fin - 1986) = 0")))$confint[1:3],
                       confint(glht(mod_1986, linfct = c("classWrks + 32*classWrks:I(int_year_fin - 1986) = 0")))$confint[1:3],
                       confint(glht(mod_1986, linfct = c("classNLFs + 32*classNLFs:I(int_year_fin - 1986) = 0")))$confint[1:3]))),
            c(length(mod_1986$residuals), NA, NA, NA)),
      col.names=c(" ", "HR", "Lower", "Upper", "HR", "Lower", "Upper", "N"),
      caption="Ref: IBOs in given year",
      digits=2) %>%
  kable_styling("striped") %>%
  add_header_above(c(" ", "HR in 1986"=3, "HR in 2018"=3, " "))
```

### Including employed "chief executives, general and operative managers, and legislators" in IBO class rather than managerial class from 2005-2018

HR similar to 2001-2019 analysis above. IBO class increases from 5282 to 7650 with change (2368 respondents move from managerial class to IBO class).

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(year>=2005) %>%
  mutate(ceo=ifelse(occ==1, 1, 0),
         class_ceo=factor(ifelse(empstat==220, "NLFs",
                                ifelse((empstat>=200 & empstat<=214) | (classwk2>=1 & classwk2<=4 & managers!=1  & ceo!=1), "Workers",
                                      ifelse(classwk2>=1 & classwk2<=4 & managers==1 & ceo!=1, "Managers",
                                            ifelse((classwk2==5 & jobsecincorp==2) | (classwk2>=1 & classwk2<=4 & managers==1 & ceo==1), "IBOs",
                                                    ifelse((classwk2==6  & ceo!=1) | (classwk2==5 & jobsecincorp==1 & ceo!=1), "UBOs", NA))))), 
                          levels=c("IBOs", "UBOs", "Managers", "Workers", "NLFs"))) -> dat_sub_no_hisp_2005_ceo

dat_sub_no_hisp_2005_ceo %>%
  filter(!is.na(class_ceo)) %>% #small amount of missingness because using alternative occupation variable
  mutate(sw_ceo=ipwpoint(exposure=class_ceo,
                         family="multinomial",
                         link="logit",
                         numerator=~1,
                         denominator=~rcs(age, 3) + sex + rcs(int_year_fin, 3), 
                         data=subset(dat_sub_no_hisp_2005_ceo, !is.na(class_ceo)),
                         weights=mortwt_f,
                         maxit=1000,
                         trace=FALSE)$ipw.weights,
         sw_f_ceo=sw_ceo*mortwt_f)  -> dat_sub_no_hisp_2005_ceo

summary(dat_sub_no_hisp_2005_ceo$sw_ceo)
```

```{r warning=FALSE, message=FALSE}
kapped_ceo <- survfit(Surv(time,dead)~class, robust=T, w=sw_f_ceo, data=dat_sub_no_hisp_2005_ceo, se=T)
mod_ceo <- coxph(Surv(time,dead)~class, robust=T, w=sw_f_ceo, data=dat_sub_no_hisp_2005_ceo)
binded_ceo <- survdiffed(kapped_ceo, timed=15)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_ceo, bind=binded_ceo)
```

# Session info

```{r  warning=FALSE, message=FALSE}
sessionInfo()
```
