---
title: "Dead labor: classed, gendered, and racialized mortality inequities in the United States"
author: "Jerzy Eisenberg-Guyot"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    keep_md: true
    toc_depth: 5
always_allow_html: true
---

* Overview: 
  + Descriptive analyses of class and mortality in the cross-sectional 1986-2014 NHIS with mortality follow-up through the end of 2015. Survival curves and Cox models are minimally adjusted (age, gender, and year) using IPW unless otherwise noted.
<br>
* Coding of classes:
  + Capitalists are those who are self-employed in an incorporated business
  + The petite bourgeoisie (PBs) are those who are self-employed in an unincorporated business or working without pay in a family business/farm (very few people in the latter)
  + Managers are those who are not self-employed but who are employed and who have an "executive, administrative, or managerial" occupation
  + Workers are those who are unemployed/"not employed" (few people) OR who are employed but do not have a management occupation
  + Not in the labor force (NILFs) are those who are "not in the labor force" 
<br>
* To-do:
  + Decide whether we want to keep worker subdivision (white collar, service, blue-collar/farming, unemployed) as a secondary analysis or incorporate it into our primary analyses. Imprecision could become problematic.
    - Based on: https://ajph.aphapublications.org/doi/pdfplus/10.2105/AJPH.2011.300173 and https://pubmed.ncbi.nlm.nih.gov/21671459/
    - Could combine services, blue collar/farming, and unemployed, given similar survival probabilities among them
<br>
* General notes:
    + We could just call capitalists/PBs incorporated business owners/unincorporated business owners if we're worried the class measure is too imprecise (and say they're analogous to capitalists/PBs in the text)
    + Numemps (number of employees at work) only available from 2001-2014 - would be alternative way to classify capitalists/PBs but I think its categories are too imprecise and limited availability is problematic
    + Can't identify "chief executives, general and operative managers, and legislators" until 2004 because of limited occupation data before then (can never identify "chief executives" alone, who we've classified as capitalists in the past)
    + Originally was hoping to examine cause-specific mortality, but I'm not sure it makes sense as counts are small, detailed cause of death not available from 2005-on, and would need to handle competing events
    + I don't think it's worth doing MICE since there's very little missingness and it'd be incredibly computationally intensive given the size of the dataset
    + Tried subdividing POC into finer racialized groups but estimates were very imprecise, particularly among NH Black and Hispanic capitalists
    + Useful citation regarding occupation and mortality in NHIS: https://www.sciencedirect.com/science/article/pii/S0091743520301535 and http://paa2019.populationassociation.org/uploads/190718
<br>
* Methods notes: 
    + When robust=T, weights are treated properly by survfit and coxph functions (as probability or sampling weights rather than frequency weights)
    + Citations for propensity score approach: https://www.ncbi.nlm.nih.gov/labs/pmc/articles/PMC5802372/ and https://journals.sagepub.com/doi/full/10.1177/0193841X20938497
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(data.table)
library(here)
library(survival)
library(rms)
library(broom)
library(Rcpp)
library(rstpm2)
library(survey)
library(kableExtra)
library(tableone)
library(survminer)
library(ipw)
library(directlabels)
library(RColorBrewer)
library(patchwork)
options(scipen=999)
options(knitr.kable.NA = '')

#load dat
dat <- fread(here('nhis_1986_2014.csv'))

#make all variable names lowercase
dat %>%
  rename_all(tolower) -> dat

#exclude those outside age ranges (age has 0.09% missing)
#exclude those ineligible for mortality follow-up and those <25 or 65+ and those who aren't sample adults from 1997-on (occ1995 and classwk not available for others after that time)
#exclude 1997-2000, when there's no data on whether business is incorporated or not
#exclude 2015-on, when there's no mortality data
#we'll exclude 1992 Hispanic oversample (only 3517 additional respondents) (per advice of IPUMS) separately, as we want to include them in sensitivity analyses
dat %>%
  filter((age>=25 & age<65) & mortelig==1 & (is.na(astatflg) | astatflg==1) & (year<1997 | (year>2000 & year<2015))) -> dat_sub

#make variables
dat_sub %>%
  mutate_at(.vars=(vars(-c(mortucod, mortwt, mortwtsa, psu, strata, nhispid, nhishid))),
            ~ifelse(. %in% c(91, 96, 97, 98, 99, 970, 980, 990, 997, 999, 7777, 8888, 9999), NA, .)) %>%
  mutate(classwk2=ifelse(classwk2 %in% (7:9), NA, classwk2), #these are real values in some other variables so we'll reset them here
         jobsecincorp=ifelse(jobsecincorp %in% (7:9), NA, jobsecincorp),
         racesr=ifelse(racesr==900, NA, racesr),
         lamtwrk=ifelse(lamtwrk %in% 7:9, NA, lamtwrk)) %>%
  mutate(hispanic_oversample=ifelse((year==1992 & substr(nhispid, 1, 4)==1991), 1, 0),
         year_1986=year-1986, #center year at beginning and end of follow-up
         year_2014=year-2014,
         mortwt_f=ifelse(year<=1996, mortwt, mortwtsa), #per advice of IPUMS help desk, use mortwt from 1986-1996 and mortwtsa from 1997-on
         dead=ifelse(mortstat==1, 1, 0),
         mortdodq=ifelse(mortdodq==1, 91.25, #assume deaths occurred at the end of the quarter
                         ifelse(mortdodq==2, 182.5,
                                ifelse(mortdodq==3, 273.75, 
                                       ifelse(mortdodq==4, 365, NA)))),
         time=ifelse(dead==0, 2016 - year, mortdody + (mortdodq/365) - year), #deaths were allowed to occur through the end of Dec 31 2015, i.e., 2016
         dead_86_95_96=ifelse(year>1995, NA,
                              ifelse(year<=1995 & (is.na(mortdody) | mortdody>1996), 0,
                                    ifelse(year<=1995 & mortdody<=1996, 1, NA))),
         time_86_95_96=ifelse(year>1995, NA,
                              ifelse(dead_86_95_96==1, mortdody + (mortdodq/365) - year,
                                     ifelse(dead_86_95_96==0, 1997-year, NA))),
         race=ifelse(racesr==100, "White", 
                     ifelse(racesr==200, "Black", "Other")),
         hispeth_bin=ifelse(hispeth!=10, 1, 0),
         race_h=factor(ifelse(hispeth!=10, "Hispanic",
                              ifelse(hispeth==10 & race=="White", "NH white",
                                     ifelse(hispeth==10 & race=="Black", "NH black", "NH other"))), levels=c("NH white", "NH black", "Hispanic", "NH other")),
         poc=ifelse(race_h=="NH white", "NH white", "POC"),
         sex=ifelse(sex==1, "male", "female"),
         race_h_sex=paste(race_h, sex),
         educ=factor(ifelse(educrec1<13, "<HS",
                            ifelse(educrec1==13, "HS",
                                   ifelse(educrec1==14, "Some college", "College+"))), levels=c("<HS", "HS", "Some college", "College+")),
         marital_tri=factor(ifelse(marstat %in% 10:12, "Married",
                                   ifelse(marstat %in% 20:40, "Widowed/divorced/separated", 
                                          ifelse(marstat==50, "Single", NA))), levels=c("Married", "Single", "Widowed/divorced/separated")),
         region=ifelse(region==1, "NE", 
                       ifelse(region==2, "MW",
                              ifelse(region==3, "S", "W"))),
         incimp1_rev=ifelse(year<2007, NA,
                            ifelse(incimp1>66, 66, incimp1)),
         managers=ifelse(occ1995>=102 & occ1995<=104, 1, 0),
         class=factor(ifelse(empstat==220, "NILFs",
                             ifelse((empstat>=200 & empstat<=214) | (year<2001 & classwk>=20 & classwk<=34 & managers!=1) | (year>=2001 & classwk2>=1 & classwk2<=4 & managers!=1), "Wrks",
                                    ifelse((year<2001 & classwk>=20 & classwk<=34 & managers==1) | (year>=2001 & classwk2>=1 & classwk2<=4 & managers==1), "Mans",
                                           ifelse((year<2001 & classwk==41) | (year>=2001 & classwk2==5 & jobsecincorp==2), "Caps",
                                                  ifelse((year<2001 & (classwk==42 | classwk==50)) | ((year>=2001 & classwk2==6) | (year>=2001 & classwk2==5 & jobsecincorp==1)), "PBs", NA))))), 
                     levels=c("Caps", "PBs", "Mans", "Wrks", "NILFs")),
         class_gender=factor(ifelse(class=="Wrks" & sex=="male", "Male wrks",
                                    ifelse(class=="Wrks" & sex=="female", "Female wrks",
                                           ifelse(class=="Mans" & sex=="male", "Male mans",
                                                  ifelse(class=="Mans" & sex=="female", "Female mans",
                                                         ifelse(class=="PBs" & sex=="male", "Male PBs",
                                                                ifelse(class=="PBs" & sex=="female", "Female PBs",
                                                                       ifelse(class=="Caps" & sex=="male", "Male caps",
                                                                              ifelse(class=="Caps" & sex=="female", "Female caps",
                                                                                     ifelse(class=="NILFs" & sex=="male", "Male NILFs",
                                                                                            ifelse(class=="NILFs" & sex=="female", "Female NILFs", NA)))))))))),
                             levels=c("Male caps", "Male PBs", "Male mans", "Male wrks", "Male NILFs",
                                      "Female caps", "Female PBs", "Female mans", "Female wrks", "Female NILFs")),
         class_poc=factor(ifelse(class=="Wrks" & poc=="NH white", "NH white wrks",
                                 ifelse(class=="Wrks" & poc=="POC", "POC wrks",
                                        ifelse(class=="Mans" & poc=="NH white", "NH white mans",
                                               ifelse(class=="Mans" & poc=="POC", "POC mans",
                                                      ifelse(class=="PBs" & poc=="NH white", "NH white PBs",
                                                             ifelse(class=="PBs" & poc=="POC", "POC PBs",
                                                                    ifelse(class=="Caps" & poc=="NH white", "NH white caps",
                                                                           ifelse(class=="Caps" & poc=="POC", "POC caps", 
                                                                                  ifelse(class=="NILFs" & poc=="NH white", "NH white NILFs", 
                                                                                         ifelse(class=="NILFs" & poc=="POC", "POC NILFs", NA)))))))))),
                          levels=c("NH white caps", "NH white PBs", "NH white mans", "NH white wrks", "NH white NILFs",
                                   "POC caps", "POC PBs", "POC mans", "POC wrks", "POC NILFs")),
         class_sens=factor(ifelse(year<2001 | year>2003, NA, #identical to class variable
                                  ifelse(empstat==220, "NILFs",
                                         ifelse((empstat>=200 & empstat<=214) | (classwk>=20 & classwk<=34 & managers!=1), "Wrks",
                                                ifelse((classwk>=20 & classwk<=34 & managers==1), "Mans",
                                                       ifelse(classwk==40 & jobsecincorp==2, "Caps",
                                                              ifelse((classwk==40 & jobsecincorp==1) | classwk==50, "PBs", NA)))))), 
                          levels=c("Caps", "PBs", "Mans", "Wrks", "NILFs")),
         class_occ=factor(ifelse(empstat==220, "NILFs",
                                 ifelse(empstat>=200 & empstat<=214, "Unemps",
                                        ifelse((year<2001 & classwk>=20 & classwk<=34 & occ1995>=900) | (year>=2001 & classwk2>=1 & classwk2<=4 & occ1995>=900), "Blue collar/farm",
                                               ifelse((year<2001 & classwk>=20 & classwk<=34 & occ1995>=600 & occ1995<900) | (year>=2001 & classwk2>=1 & classwk2<=4 & occ1995>=600 & occ1995<900), "Services",
                                                      ifelse((year<2001 & classwk>=20 & classwk<=34 & occ1995>=200 & occ1995<600) | (year>=2001 & classwk2>=1 & classwk2<=4 & occ1995>=200 & occ1995<600), "White collar",
                                                             ifelse((year<2001 & classwk>=20 & classwk<=34 & managers==1) | (year>=2001 & classwk2>=1 & classwk2<=4 & managers==1), "Mans",
                                                                    ifelse((year<2001 & classwk==41) | (year>=2001 & classwk2==5 & jobsecincorp==2), "Caps",
                                                                           ifelse((year<2001 & (classwk==42 | classwk==50)) | ((year>=2001 & classwk2==6) | (year>=2001 & classwk2==5 & jobsecincorp==1)), "PBs", NA)))))))), 
                     levels=c("Caps", "PBs", "Mans", "White collar", "Services", "Blue collar/farm", "Unemps", "NILFs"))) -> dat_sub
                                                                                     
#exclude couple respondents with negative follow-up time
dat_sub %>%
  filter(time>=0) -> dat_sub

#make dataset without hispanic oversample for main analyses
dat_sub %>%
  filter(hispanic_oversample==0) -> dat_sub_no_hisp
```

# Missingness 

Age missingness is in full dataset among years and samples of interest; missingness for other variables is among those ages 25-64

```{r message=FALSE, warning=FALSE}
#age in full dataset
dat %>%
  filter(mortelig==1 & (is.na(astatflg) | astatflg==1) & (year<1997 | (year>2000 & year<2015)) & 
           !(year==1992 & substr(nhispid, 1, 4)==1991)) %>%
  mutate_at(.vars=(vars(-c(mortucod, mortwt, mortwtsa, psu, strata, nhispid, nhishid))), ~ifelse(. %in% c(91, 96, 97, 98, 99, 970, 980, 990, 7777, 8888, 9999), NA, .)) %>%
  summarise_at(vars('age'), funs(na=100*sum(is.na(.)/1233802))) %>%
  mutate(name='age') %>% 
  relocate(na, .after=last_col()) -> dat_age_na

#other vars in subsetted dataset
dat_sub_no_hisp %>%
  summarise_at(vars(c('class', 'sex', 'race_h', 'educ', 'marital_tri', 'region')), funs(100*sum(is.na(.)/861724))) %>%
  tidyr::pivot_longer(class:region, values_to='na') -> dat_other_na

kable(bind_rows(dat_age_na, dat_other_na), digits=2, col.names=c("Variable", "Percent missing")) %>%
  kable_styling('striped')
```

# Descriptives

Excluding missingness unless otherwise noted

## Unweighted and stratified table one with missingness in variables of interest (aside from age and class)

```{r message=FALSE, warning=FALSE}
#vars of interest
vars <- c('sex', 'race_h', 'educ', 'marital_tri', 'region', 'age', 'dead')
catvars <- c("sex", 'dead')
nonorm <- c('age')

x <- CreateTableOne(data = dat_sub_no_hisp, vars=vars, factorVars=catvars, strata='class', includeNA=TRUE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, test=FALSE)
kable(x) %>%
  kable_styling(c("striped", "condensed"))
```

## Weighted and stratified table one 

```{r message=FALSE, warning=FALSE}
dat_sub_no_hisp_svy <- svydesign(ids = ~ psu,
                         strata = ~ strata, 
                         weights = ~ mortwt_f,
                         nest=TRUE, 
                         data=dat_sub_no_hisp)

x <- svyCreateTableOne(data = dat_sub_no_hisp_svy, vars=vars, factorVars=catvars, strata='class', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
kable(x) %>%
  kable_styling(c("striped", "condensed"))
```

## Weighted and stratified class composition of gender-races 

```{r message=FALSE, warning=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, !is.na(race_h)), vars='class', factorVars='class', strata='race_h_sex', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
kable(x[2:7,]) %>%
  kable_styling(c("striped", "condensed"))
```

## Weighted proportions in classes over time 

```{r message=FALSE, warning=FALSE}
dat_sub_no_hisp %>%
  filter(!is.na(class)) %>%
  group_by(class, year) %>%
  summarise(n = sum(mortwt_f)) %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(prop = n / sum(n)) -> propped

ggplot(propped, aes(x=year, y=prop, group=class, color=class, label=class)) +
  geom_dl(method='last.qp') +
  geom_line() +
  scale_color_manual(values=brewer.pal(n = 12, name = "Paired")[c(2,8,4,10,6)]) +
  scale_x_continuous(expand=expansion(mult=c(0,0.08))) +
  scale_y_continuous(limits=c(0, 0.63), expand=expansion(mult=c(0,0))) +
  xlab("Year") +
  ylab("Proportion") +
  theme_light() +
  theme(legend.position = "none")
```

## Family income distribution across classes from 2007-2014 

Income coding inconsistent prior to 2007. Unadjusted for inflation. Row label shows lower bound of category.

```{r message=FALSE, warning=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, year>=2007), vars='incimp1_rev', factorVars='incimp1_rev', strata='class', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
row.names(x)[3:23] <- c(paste0(seq(0,95,5), "k"),">=100k")
kable(x[3:23,]) %>%
  kable_styling(c("striped", "condensed"))
```

## Number of employees at work across classes from 2001-2014

Excluding those not working. Capitalists 4.4x as likely as PBs to work in firms with 10+ employees. It probably makes sense that managers and workers are much more likely than capitalists to work in larger firms, since any given firm, regardless of size, will only have one capitalist (in theory), whereas such firms will have multiple workers. For example, in a society with 5 firms, 4 of which have 5 employees and 1 of which has 1000 employees, 80% of capitalists will come from firms with 5 employees, but 98% (1000/1020) of workers will come from firms with 1000+ employees.

```{r message=FALSE, warning=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, year>=2001 & numemps!=0 & class!="NILFs"), vars='numemps', factorVars='numemps', strata='class', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
row.names(x)[3:10] <- c("1-9", "10-24", "25-49", "50-99", "100-249", "250-499", "500-999", "1000+")
kable(x[3:10,1:4]) %>%
  kable_styling(c("striped", "condensed"))
```

## Reason not in labor force

### Main reason for not working last week

Variable only available from 2004-on.

```{r warning=FALSE, message=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, year>=2004 & empstat==220 & whynowk2!=0), vars='whynowk2', factorVars='whynowk2', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
row.names(x)[3:12] <- c("Keeping house", "Going to school", "Retired", "Unable to work for health reasons", "Disabled", "On layoff", "On planned vacation", "On family leave", "Have job/contract; off season", "Other")
kable(x[3:12,], digits=2, col.names="Percent") %>% 
  kable_styling("striped")
```

### Limited in kind/amount of work can do

Variable only available from 1997-on.

```{r warning=FALSE, message=FALSE}
x <- svyCreateTableOne(data = subset(dat_sub_no_hisp_svy, year>=1997), vars='lamtwrk', factorVars='lamtwrk', strata='empstat', includeNA=FALSE)
x <- print(x, printToggle=FALSE, noSpaces=TRUE, nonnormal=nonorm, format='p', test=FALSE)
row.names(x)[3:5] <- c("Not limited", "Limited", "Unable to work")
kable(x[3:5,], col.names=c("Working for pay", "Working w/o pay", "With job, but not at work", "Not employed", "NILF")) %>% 
  kable_styling("striped")
```

# Class and all-cause mortality

```{r}
plotted <- function(datted=kapped, bottom=0.60){
  
  survplot <- ggsurvplot(datted,
                         size=0.35,
                         conf.int = T,
                         ylim=c(bottom, 1.0),
                         xlim=c(0, max(datted$time)),
                         axes.offset=F,
                         censor=F)
  survplot$plot +
    scale_color_manual(values=brewer.pal(n = 12, name = "Paired")[c(2,8,4,10,6)]) +
    scale_fill_manual(values=brewer.pal(n = 12, name = "Paired")[c(2,8,4,10,6)]) +
    geom_dl(aes(label=class, group=strata, color=strata), method='last.qp') +
    scale_x_continuous(expand=expansion(mult=c(0,0.09))) +
    scale_y_continuous(limits=c(bottom, 1.0), expand=expansion(mult=c(0,0))) +
    ylab("Probability of survival") +
    xlab("Year") +
    theme_light() +
    theme(legend.position="none")

}

#survdiff function
survdiffed <- function(datted=kapped, timed=30){
  
  #end of f/u
  point <- summary(datted,times=timed)
  
  #pb-cap
  pb <- 100*(point$surv[2] - point$surv[1])
  pb_lower <- 100*(pb/100 - 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  pb_upper <- 100*(pb/100 + 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
    
  #mc-cap
  mc <- 100*(point$surv[3] - point$surv[1])
  mc_lower <- 100*(mc/100 - 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  mc_upper <- 100*(mc/100 + 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  
  #worker-cap
  wc <- 100*(point$surv[4] - point$surv[1])
  wc_lower <- 100*(wc/100 - 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  wc_upper <- 100*(wc/100 + 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  
  #nilf-cap
  nc <- 100*(point$surv[5] - point$surv[1])
  nc_lower <- 100*(nc/100 - 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  nc_upper <- 100*(nc/100 + 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  
  rbind(c(pb, pb_lower, pb_upper), c(mc, mc_lower, mc_upper), c(wc, wc_lower, wc_upper), c(nc, nc_lower, nc_upper))
  
}

#survdiff function for groups
survdiffed_group <- function(datted=kapped, timed=30){
  
  #end of f/u
  point <- summary(datted,times=timed)
  
  #pb-cap
  pb <- 100*(point$surv[2] - point$surv[1])
  pb_lower <- 100*(pb/100 - 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  pb_upper <- 100*(pb/100 + 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  
  #mc-cap
  mc <- 100*(point$surv[3] - point$surv[1])
  mc_lower <- 100*(mc/100 - 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  mc_upper <- 100*(mc/100 + 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  
  #worker-cap
  wc <- 100*(point$surv[4] - point$surv[1])
  wc_lower <- 100*(wc/100 - 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  wc_upper <- 100*(wc/100 + 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  
  #nilf-cap
  nc <- 100*(point$surv[5] - point$surv[1])
  nc_lower <- 100*(nc/100 - 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  nc_upper <- 100*(nc/100 + 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  
  #cap2-cap
  cap2 <- 100*(point$surv[6] - point$surv[1])
  cap2_lower <- 100*(cap2/100 - 1.96*sqrt(point$std.err[6]^2 + point$std.err[1]^2))
  cap2_upper <- 100*(cap2/100 + 1.96*sqrt(point$std.err[6]^2 + point$std.err[1]^2))
  
  #pb2-cap
  pb2 <- 100*(point$surv[7] - point$surv[1])
  pb2_lower <- 100*(pb2/100 - 1.96*sqrt(point$std.err[7]^2 + point$std.err[1]^2))
  pb2_upper <- 100*(pb2/100 + 1.96*sqrt(point$std.err[7]^2 + point$std.err[1]^2))
  
  #man2-cap
  man2 <- 100*(point$surv[8] - point$surv[1])
  man2_lower <- 100*(man2/100 - 1.96*sqrt(point$std.err[8]^2 + point$std.err[1]^2))
  man2_upper <- 100*(man2/100 + 1.96*sqrt(point$std.err[8]^2 + point$std.err[1]^2))
  
  #wrk2-cap
  wrk2 <- 100*(point$surv[9] - point$surv[1])
  wrk2_lower <- 100*(wrk2/100 - 1.96*sqrt(point$std.err[9]^2 + point$std.err[1]^2))
  wrk2_upper <- 100*(wrk2/100 + 1.96*sqrt(point$std.err[9]^2 + point$std.err[1]^2))
  
  #nilf2-cap
  nilf2 <- 100*(point$surv[10] - point$surv[1])
  nilf2_lower <- 100*(nilf2/100 - 1.96*sqrt(point$std.err[10]^2 + point$std.err[1]^2))
  nilf2_upper <- 100*(nilf2/100 + 1.96*sqrt(point$std.err[10]^2 + point$std.err[1]^2))
  
  rbind(c(pb, pb_lower, pb_upper), c(mc, mc_lower, mc_upper), c(wc, wc_lower, wc_upper), c(nc, nc_lower, nc_upper),
        c(cap2, cap2_lower, cap2_upper), c(pb2, pb2_lower, pb2_upper), c(man2, man2_lower, man2_upper), c(wrk2, wrk2_lower, wrk2_upper), c(nilf2, nilf2_lower, nilf2_upper))
  
}

#survdiff function for worker subdivision
survdiffed_sub <- function(datted=kapped, timed=30){
  
  #end of f/u
  point <- summary(datted,times=timed)
  
  #pb-cap
  pb <- 100*(point$surv[2] - point$surv[1])
  pb_lower <- 100*(pb/100 - 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  pb_upper <- 100*(pb/100 + 1.96*sqrt(point$std.err[2]^2 + point$std.err[1]^2))
  
  #mc-cap
  mc <- 100*(point$surv[3] - point$surv[1])
  mc_lower <- 100*(mc/100 - 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  mc_upper <- 100*(mc/100 + 1.96*sqrt(point$std.err[3]^2 + point$std.err[1]^2))
  
  #whitecollar-cap
  wc <- 100*(point$surv[4] - point$surv[1])
  wc_lower <- 100*(wc/100 - 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  wc_upper <- 100*(wc/100 + 1.96*sqrt(point$std.err[4]^2 + point$std.err[1]^2))
  
  #services-cap
  servs <- 100*(point$surv[5] - point$surv[1])
  servs_lower <- 100*(servs/100 - 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  servs_upper <- 100*(servs/100 + 1.96*sqrt(point$std.err[5]^2 + point$std.err[1]^2))
  
  #bluecollar-cap
  bc <- 100*(point$surv[6] - point$surv[1])
  bc_lower <- 100*(bc/100 - 1.96*sqrt(point$std.err[6]^2 + point$std.err[1]^2))
  bc_upper <- 100*(bc/100 + 1.96*sqrt(point$std.err[6]^2 + point$std.err[1]^2))
  
  #pb2-cap
  unemp <- 100*(point$surv[7] - point$surv[1])
  unemp_lower <- 100*(unemp/100 - 1.96*sqrt(point$std.err[7]^2 + point$std.err[1]^2))
  unemp_upper <- 100*(unemp/100 + 1.96*sqrt(point$std.err[7]^2 + point$std.err[1]^2))
  
  #man2-cap
  nilf <- 100*(point$surv[8] - point$surv[1])
  nilf_lower <- 100*(nilf/100 - 1.96*sqrt(point$std.err[8]^2 + point$std.err[1]^2))
  nilf_upper <- 100*(nilf/100 + 1.96*sqrt(point$std.err[8]^2 + point$std.err[1]^2))
  
  rbind(c(pb, pb_lower, pb_upper), c(mc, mc_lower, mc_upper), c(wc, wc_lower, wc_upper), c(servs, servs_lower, servs_upper),
        c(bc, bc_lower, bc_upper), c(unemp, unemp_lower, unemp_upper), c(nilf, nilf_lower, nilf_upper))
  
}

#tidying regression output function
tidy_n <- function(modded=mod, rows=1:4, cols=c(1,2,7,8), bind=binded, nad=3, captioned="Ref: capitalist") {
  tidydf <- cbind(tidy(modded, exponentiate=T, conf.int=T)[rows, cols], bind)
  tidydf$N <- c(length(modded$residuals), rep(NA, nad))
  kable(tidydf, digits=c(2,2,2,2,1,1,1,1), caption=captioned, col.names=c(" ", "HR", "Lower", "Upper", "SD", "Lower", "Upper", "N")) %>%
    kable_styling("striped") %>%
    add_header_above(c(" ", "HR"=3, "SD per 100 at end of f/u"=3, " "))
}
```

## Overall

### Less-adjusted (age, gender, year)

#### Distribution of IPW

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(!is.na(class) & !is.na(age)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 5), 
                     data = subset(dat_sub_no_hisp, !is.na(class) & !is.na(age)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_overall

summary(dat_sub_no_hisp_overall$sw)
```

```{r warning=FALSE, message=FALSE}
kapped <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_overall, se=T)
mod <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_overall)
binded <- survdiffed()
```

#### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=6}
plotted()

ggsave("overall_survival.png", dpi=600, height=4, width=6)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n()
```

### More-adjusted (age, gender, year, education, marital status, region, and racialized group)

#### Distribution of IPW

Truncated at 0.5% and 99.5%.

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(!is.na(class) & !is.na(age) & !is.na(educ) & !is.na(marital_tri) & !is.na(region) & !is.na(race_h)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 5) + educ + marital_tri + region + race_h, 
                     data = subset(dat_sub_no_hisp, !is.na(class) & !is.na(age) & !is.na(educ) & !is.na(marital_tri) & !is.na(region) & !is.na(race_h)),
                     weights=mortwt_f,
                     trace=FALSE,
                     trunc=0.005)$weights.trunc,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_overall_adj

summary(dat_sub_no_hisp_overall_adj$sw)
```

```{r warning=FALSE, message=FALSE}
kapped <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_overall_adj, se=T)
mod <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_overall_adj)
binded <- survdiffed()
```

#### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=6}
plotted()

ggsave("overall_survival_adj.png", dpi=600, height=4, width=6)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n()
```

## Change over time - 1986-1995 waves w/ f/u until end of 1996 vs 2005-2014 waves w/ f/u through end of 2015

### Less-adjusted

#### Distribution of IPW

##### 1986-1995

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(year<=1995 & !is.na(class) & !is.na(age)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 3), 
                     data = subset(dat_sub_no_hisp, year<=1995 & !is.na(class) & !is.na(age)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f,
         period="1986-1995 with follow-up through 1996")  -> dat_sub_no_hisp_1986

summary(dat_sub_no_hisp_1986$sw)
```

##### 2005-2014

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(year>=2005 & !is.na(class) & !is.na(age)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 3), 
                     data = subset(dat_sub_no_hisp, year>=2005 & !is.na(class) & !is.na(age)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f,
         period="2005-2014 with follow-up through 2015")  -> dat_sub_no_hisp_2005

summary(dat_sub_no_hisp_2005$sw)
```

```{r warning=FALSE, message=FALSE}
#early
kapped_early <- survfit(Surv(time_86_95_96, dead_86_95_96)~class + period, robust=T, w=sw_f, data=dat_sub_no_hisp_1986, se=T) #period allows for faceting but doesnt affect ests
mod_early <- coxph(Surv(time_86_95_96, dead_86_95_96)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_1986)
binded_early <- survdiffed(kapped_early, timed=11)

#late
kapped_late <- survfit(Surv(time,dead)~class + period, robust=T, w=sw_f, data=dat_sub_no_hisp_2005, se=T)
mod_late <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_2005)
binded_late <- survdiffed(kapped_late, timed=11)
```

#### IPW survival plots

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.5}
plotted(kapped_early) +  
  facet_wrap(~period) + 
  scale_y_continuous(limits=c(0.88, 1.0), breaks=c(0.88, 0.91, 0.94, 0.97, 1.0), expand=expansion(mult=c(0,0))) +
  scale_x_continuous(limits=c(0, 11), breaks=seq(1, 11, 1), expand=expansion(mult=c(0,0.13))) +
  theme(panel.grid.minor.x = element_blank(), strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black")) +
plotted(kapped_late) +  
  facet_wrap(~period) + 
  scale_y_continuous(limits=c(0.88, 1.0), breaks=c(0.88, 0.91, 0.94, 0.97, 1.0), expand=expansion(mult=c(0,0))) +
  scale_x_continuous(limits=c(0, 11), breaks=seq(1, 11, 1), expand=expansion(mult=c(0,0.13))) +
  theme(panel.grid.minor.x = element_blank(), strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black"),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())

ggsave("change_over_time.png", dpi=600, height=4, width=8.5)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_early, bind=binded_early, captioned="Ref: capitalists (1986-1995 subset)")
tidy_n(modded=mod_late, bind=binded_late, captioned="Ref: capitalists (2005-2014 subset)")
```

### More-adjusted

#### Distribution of IPW

Truncated at 0.5% and 99.5%.

##### 1986-1995

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(year<=1995 & !is.na(class) & !is.na(age) & !is.na(educ) & !is.na(marital_tri) & !is.na(region) & !is.na(race_h)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 5) + educ + marital_tri + region + race_h, 
                     data = subset(dat_sub_no_hisp, year<=1995 & !is.na(class) & !is.na(age) & !is.na(educ) & !is.na(marital_tri) & !is.na(region) & !is.na(race_h)),
                     weights=mortwt_f,
                     trace=FALSE,
                     trunc=0.005)$weights.trunc,
         sw_f=sw*mortwt_f,
         period="1986-1995 with follow-up through 1996")  -> dat_sub_no_hisp_1986

summary(dat_sub_no_hisp_1986$sw)
```

##### 2005-2014

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(year>=2005 & !is.na(class) & !is.na(age) & !is.na(educ) & !is.na(marital_tri) & !is.na(region) & !is.na(race_h)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 5) + educ + marital_tri + region + race_h, 
                     data = subset(dat_sub_no_hisp, year>=2005 & !is.na(class) & !is.na(age) & !is.na(educ) & !is.na(marital_tri) & !is.na(region) & !is.na(race_h)),
                     trace=FALSE,
                     trunc=0.005)$weights.trunc,
         sw_f=sw*mortwt_f,
         period="2005-2014 with follow-up through 2015")  -> dat_sub_no_hisp_2005

summary(dat_sub_no_hisp_2005$sw)
```

```{r warning=FALSE, message=FALSE}
#early
kapped_early <- survfit(Surv(time_86_95_96, dead_86_95_96)~class + period, robust=T, w=sw_f, data=dat_sub_no_hisp_1986, se=T) #period allows for faceting but doesnt affect ests
mod_early <- coxph(Surv(time_86_95_96, dead_86_95_96)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_1986)
binded_early <- survdiffed(kapped_early, timed=11)

#late
kapped_late <- survfit(Surv(time,dead)~class + period, robust=T, w=sw_f, data=dat_sub_no_hisp_2005, se=T)
mod_late <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_2005)
binded_late <- survdiffed(kapped_late, timed=11)
```

#### IPW survival plots

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.5}
plotted(kapped_early) +  
  facet_wrap(~period) + 
  scale_y_continuous(limits=c(0.88, 1.0), breaks=c(0.88, 0.91, 0.94, 0.97, 1.0), expand=expansion(mult=c(0,0))) +
  scale_x_continuous(limits=c(0, 11), breaks=seq(1, 11, 1), expand=expansion(mult=c(0,0.13))) +
  theme(panel.grid.minor.x = element_blank(), strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black")) +
plotted(kapped_late) +  
  facet_wrap(~period) + 
  scale_y_continuous(limits=c(0.88, 1.0), breaks=c(0.88, 0.91, 0.94, 0.97, 1.0), expand=expansion(mult=c(0,0))) +
  scale_x_continuous(limits=c(0, 11), breaks=seq(1, 11, 1), expand=expansion(mult=c(0,0.13))) +
  theme(panel.grid.minor.x = element_blank(), strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black"),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())

ggsave("change_over_time_adj.png", dpi=600, height=4, width=8.5)
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n(modded=mod_early, bind=binded_early, captioned="Ref: capitalists (1986-1995 subset)")
tidy_n(modded=mod_late, bind=binded_late, captioned="Ref: capitalists (2005-2014 subset)")
```

## Class-by-gender interaction

### Distribution of IPW

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(!is.na(class_gender) & !is.na(age)) %>%
  mutate(sw=ipwpoint(exposure=class_gender,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + rcs(year, 3), 
                     data = subset(dat_sub_no_hisp, !is.na(class_gender) & !is.na(age)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f,
         gender=ifelse(sex=="female", "Women", "Men"))  -> dat_sub_no_hisp_gender

summary(dat_sub_no_hisp_gender$sw)
```

```{r warning=FALSE, message=FALSE}
#easier for plotting purposes to make separate fits for each gender
kapped <- survfit(Surv(time,dead)~class_gender, robust=T, w=sw_f, data=dat_sub_no_hisp_gender, se=T) 
kapped_men <- survfit(Surv(time,dead)~class + gender, robust=T, w=sw_f, data=subset(dat_sub_no_hisp_gender, gender=="Men"), se=T) 
kapped_women <- survfit(Surv(time,dead)~class + gender, robust=T, w=sw_f, data=subset(dat_sub_no_hisp_gender, gender=="Women"), se=T)

mod <- coxph(Surv(time,dead)~class_gender, robust=T, w=sw_f, data=dat_sub_no_hisp_gender)
binded <- survdiffed_group()
```

### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.5}
plotted(kapped_men, bottom=0.5) +  
  facet_wrap(~gender) + 
  scale_x_continuous(expand=expansion(mult=c(0,0.13))) +
  theme(panel.grid.minor.x = element_blank(), strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black")) +
plotted(kapped_women, bottom=0.5) +  
  facet_wrap(~gender) + 
  scale_x_continuous(expand=expansion(mult=c(0,0.13))) +
  theme(panel.grid.minor.x = element_blank(), strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black"),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())

ggsave("gender_survival.png", dpi=600, height=4, width=8.5)
```

### IPW survival difference and hazard ratio

```{r}
tidy_n(rows=1:9, nad=8, captioned="Ref: male capitalist")
```

## Class-by-racialized-group interaction

### Distribution of IPW

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(!is.na(class_poc) & !is.na(age)) %>%
  mutate(sw=ipwpoint(exposure=class_poc,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + rcs(year, 3), 
                     data = subset(dat_sub_no_hisp, !is.na(class_poc) & !is.na(age)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_race

summary(dat_sub_no_hisp_race$sw)
```

```{r warning=FALSE, message=FALSE}
#easier for plotting purposes to make separate fits for each racialized group
kapped <- survfit(Surv(time,dead)~class_poc, robust=T, w=sw_f, data=dat_sub_no_hisp_race, se=T)
kapped_white <- survfit(Surv(time,dead)~class + poc, robust=T, w=sw_f, data=subset(dat_sub_no_hisp_race, poc=="NH white"), se=T) 
kapped_poc <- survfit(Surv(time,dead)~class + poc, robust=T, w=sw_f, data=subset(dat_sub_no_hisp_race,  poc=="POC"), se=T)

mod <- coxph(Surv(time,dead)~class_poc, robust=T, w=sw_f, data=dat_sub_no_hisp_race)
binded <- survdiffed_group()
```

### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=8.5}
plotted(kapped_white) +  
  facet_wrap(~poc) + 
  scale_x_continuous(expand=expansion(mult=c(0,0.13))) +
  theme(panel.grid.minor.x = element_blank(), strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black")) +
plotted(kapped_poc) +  
  facet_wrap(~poc) + 
  scale_x_continuous(expand=expansion(mult=c(0,0.13))) +
  theme(panel.grid.minor.x = element_blank(), strip.background=element_rect(color="darkgrey", fill=NA), strip.text=element_text(color="black"),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())

ggsave("race_survival.png", dpi=600, height=4, width=8.5)
```

### IPW survival difference and hazard ratio

```{r}
tidy_n(rows=1:9, nad=8, captioned="Ref: white capitalist")
```

## Subdividing workers

### Distribution of IPW

Truncated at 0.5% and 99.5%.

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(!is.na(class_occ) & !is.na(age)) %>%
  mutate(sw=ipwpoint(exposure=class_occ,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 5), 
                     data = subset(dat_sub_no_hisp, !is.na(class_occ) & !is.na(age)),
                     weights=mortwt_f,
                     trunc=0.005,
                     trace=FALSE)$weights.trunc,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_class_occ

summary(dat_sub_no_hisp_class_occ$sw)
```

```{r warning=FALSE, message=FALSE}
kapped <- survfit(Surv(time,dead)~class_occ, robust=T, w=sw_f, data=dat_sub_no_hisp_class_occ, se=T)
mod <- coxph(Surv(time,dead)~class_occ, robust=T, w=sw_f, data=dat_sub_no_hisp_class_occ)
binded <- survdiffed_sub()
```

### IPW survival plot

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=6.5}
survplot <- ggsurvplot(kapped,
                       size=0.35,
                       conf.int = T,
                       ylim=c(0.50, 1.0),
                       xlim=c(0, max(kapped$time)),
                       axes.offset=F,
                       censor=F)

survplot$plot +
  scale_color_manual(values=c(brewer.pal(n = 12, name = "Paired")[c(2,8,4)], 
                              brewer.pal(n=9, name="Purples")[c(5,6,7,8)], brewer.pal(n = 12, name = "Paired")[c(6)])) +
  scale_fill_manual(values=c(brewer.pal(n = 12, name = "Paired")[c(2,8,4)], 
                             brewer.pal(n=9, name="Purples")[c(5,6,7,8)], brewer.pal(n = 12, name = "Paired")[c(6)])) +
  geom_dl(aes(label=class_occ, group=strata, color=strata), method='last.qp') +
  scale_x_continuous(expand=expansion(mult=c(0,0.25))) +
  scale_y_continuous(limits=c(0.6, 1.0), expand=expansion(mult=c(0,0))) +
  ylab("Probability of survival") +
  xlab("Year") +
  theme_light() +
  theme(legend.position="none")

ggsave("subdivided_workers_survival.png", dpi=600, height=4, width=6)
```

### IPW survival difference and hazard ratio

```{r}
tidy_n(rows=1:7, nad=6, captioned="Ref: capitalist")
```

## Sensitivity analyses

### Regression adjustment instead of IPW for minimally-adjusted overall Cox analyses

Similar to primary but slightly larger effect sizes for workers and PBs.

```{r warning=FALSE, message=FALSE}
mod <- coxph(Surv(time,dead)~class + rcs(age, 3) + sex + rcs(year, 5), robust=T, w=mortwt_f, data=dat_sub_no_hisp_overall)

kable(cbind(tidy(mod, exponentiate=T, conf.int=T)[1:4, c(1,2,7,8)], c(length(mod$residuals), NA, NA, NA)), 
      col.names=c(" ", "HR", "Lower", "Upper", "N"), caption="Ref: capitalist", digits=2) %>%
  kable_styling("striped")
```

### Confirming SEs make sense

#### Survey design for overall Cox analyses with IPW

Basically identical to primary. It makes no difference that we subsetted the dataset by age above rather than within the svycoxph function because age subpopulations appear in all the clusters

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp_svy <- svydesign(ids = ~ psu,
                         strata = ~ strata, 
                         weights = ~ sw_f,
                         nest=TRUE, 
                         data=dat_sub_no_hisp_overall)

mod <- svycoxph(Surv(time,dead)~class, design=subset(dat_sub_no_hisp_svy, !is.na(class) & !is.na(age)))

kable(cbind(tidy(mod, exponentiate=T, conf.int=T)[1:4, c(1,2,7,8)], c(length(mod$residuals), NA, NA, NA)), 
      col.names=c(" ", "HR", "Lower", "Upper", "N"), caption="Ref: capitalist", digits=2) %>%
  kable_styling("striped")
```

#### Clustering SEs at household level

Also basically identical to primary.

```{r warning=FALSE, message=FALSE}
kapped <- survfit(Surv(time,dead)~class + cluster(nhishid), w=sw_f, data=dat_sub_no_hisp_overall, se=T)
mod <- coxph(Surv(time,dead)~class + cluster(nhishid), w=sw_f, data=dat_sub_no_hisp_overall)
binded <- survdiffed()
```

```{r}
tidy_n()
```

### Including Hispanic oversample to see effects on point estimates 

Basically identical to primary.

#### Distribution of IPW

```{r warning=FALSE, message=FALSE}
dat_sub %>%
  filter(!is.na(class) & !is.na(age)) %>%
  mutate(sw=ipwpoint(exposure=class,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 5), 
                     data = subset(dat_sub, !is.na(class) & !is.na(age)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_hisp_overall

summary(dat_sub_hisp_overall$sw)
```

```{r warning=FALSE, message=FALSE}
kapped <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_hisp_overall, se=T)
mod <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_hisp_overall)
binded <- survdiffed()
```

#### IPW survival difference and hazard ratio

```{r}
tidy_n()
```

### Class by year interaction with regression adjustment instead of period stratification

Similar to primary.

#### Less-adjusted

```{r warning=FALSE, message=FALSE}
mod_1986 <- coxph(Surv(time,dead)~class*I(year - 1986) + rcs(age, 3) + sex, robust=T, w=mortwt_f, data=dat_sub_no_hisp_overall)
mod_2014 <- coxph(Surv(time,dead)~class*I(year - 2014) + rcs(age, 3) + sex, robust=T, w=mortwt_f, data=dat_sub_no_hisp_overall)

kable(cbind(tidy(mod_1986, exponentiate=T, conf.int=T)[1:4, c(1,2,7,8)], 
            tidy(mod_2014, exponentiate=T, conf.int=T)[1:4, c(2,7,8)],
            c(length(mod_1986$residuals), NA, NA, NA)),
            col.names=c(" ", "HR", "Lower", "Upper", "HR", "Lower", "Upper", "N"),
      caption="Ref: capitalist in given year",
      digits=2) %>%
  kable_styling("striped") %>%
  add_header_above(c(" ", "HR in 1986"=3, "HR in 2014"=3, " "))
```

#### More-adjusted

```{r warning=FALSE, message=FALSE}
mod_1986 <- coxph(Surv(time,dead)~class*I(year - 1986) + rcs(age, 3) + sex + educ + marital_tri + region + race_h, robust=T, w=mortwt_f, data=dat_sub_no_hisp_overall)
mod_2014 <- coxph(Surv(time,dead)~class*I(year - 2014) + rcs(age, 3) + sex + educ + marital_tri + region + race_h, robust=T, w=mortwt_f, data=dat_sub_no_hisp_overall)

kable(cbind(tidy(mod_1986, exponentiate=T, conf.int=T)[1:4, c(1,2,7,8)], 
            tidy(mod_2014, exponentiate=T, conf.int=T)[1:4, c(2,7,8)],-
            c(length(mod_1986$residuals), NA, NA, NA)),
            col.names=c(" ", "HR", "Lower", "Upper", "HR", "Lower", "Upper", "N"),
      caption="Ref: capitalist in given year",
      digits=2) %>%
  kable_styling("striped") %>%
  add_header_above(c(" ", "HR in 1986"=3, "HR in 2014"=3, " "))
```

### Including employed "chief executives, general and operative managers, and legislators" in capitalist class rather than managerial class from 2005-2014

Results nearly identical to 2005-2014 analysis above.

```{r warning=FALSE, message=FALSE}
dat_sub_no_hisp %>%
  filter(year>=2005) %>%
  mutate(ceo=ifelse(occ==1, 1, 0),
         class_ceo=factor(ifelse(empstat==220, "NILFs",
                                ifelse((empstat>=200 & empstat<=214) | (classwk2>=1 & classwk2<=4 & managers!=1  & ceo!=1), "Workers",
                                      ifelse(classwk2>=1 & classwk2<=4 & managers==1 & ceo!=1, "Managers",
                                            ifelse((classwk2==5 & jobsecincorp==2) | (classwk2>=1 & classwk2<=4 & managers==1 & ceo==1), "Capitalists",
                                                    ifelse((classwk2==6  & ceo!=1) | (classwk2==5 & jobsecincorp==1 & ceo!=1), "PBs", NA))))), 
                          levels=c("Capitalists", "PBs", "Managers", "Workers", "NILFs"))) -> dat_sub_no_hisp_2005_ceo

dat_sub_no_hisp_2005_ceo %>%
  filter(!is.na(class_ceo) & !is.na(age)) %>%
  mutate(sw=ipwpoint(exposure=class_ceo,
                     family="multinomial",
                     link="logit",
                     numerator=~1,
                     denominator=~rcs(age, 3) + sex + rcs(year, 3), 
                     data = subset(dat_sub_no_hisp_2005_ceo, !is.na(class_ceo) & !is.na(age)),
                     weights=mortwt_f,
                     trace=FALSE)$ipw.weights,
         sw_f=sw*mortwt_f)  -> dat_sub_no_hisp_2005_ceo

summary(dat_sub_no_hisp_2005_ceo$sw)
```

```{r warning=FALSE, message=FALSE}
kapped <- survfit(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_2005_ceo, se=T)
mod <- coxph(Surv(time,dead)~class, robust=T, w=sw_f, data=dat_sub_no_hisp_2005_ceo)
binded <- survdiffed(timed=11)
```

### IPW survival difference and hazard ratio

```{r}
tidy_n()
```

# Session info

```{r  warning=FALSE, message=FALSE}
sessionInfo()
```
